---
title: "Sourcebook > Homeownership > Mortgages"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```

    VVV EVERYTHING BELOW GOES ON PAGE VVV

# Home mortgages

The 1975 Home Mortgage Disclosure Act (HMDA) requires financial
institutions to maintain and annually disclose data about home purchases
and other loan activities. The primary purpose of this disclosure is to
determine if financial institutions are serving the housing needs of
communities and in a non-discriminatory manner. [HMDA
data](https://www.consumerfinance.gov/data-research/hmda/) help provide
an understanding of home purchase activity and allows us to look at home
loans by race and ethnicity.

    [add stock photo?]
<br> 
<br>

## Total loans by race and ethnicity

This dashboard shows the total number of loans
[originated](https://www.investopedia.com/terms/o/origination.asp) for
the purchase of a home as a primary residence.

Data is currently provided from 2018 to 2020 for all loans in Virginia
at the state, CBSA, and locality level. <br> <br>

```{r loans-race}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Homeloansbyraceandethnicity/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br> <br>
The Equal Credit Opportunity Act of 1974 prohibits lenders from discriminating on the basis of race, color, religion, national origin, sex, marital status, or age in credit transactions. As part of this legislation, race and ethnicity are voluntary information in credit applications. While this can help protect applicants from discrimination, it can make it doubly hard to understand trends in credit applications by race and ethnicity.

In some instances, a person may list their race but not their ethnicity or vice versa. This can cause under- or overcounting in the data. In other cases, a large number of applications where no race and ethnicity data was recorded can leave us with an incomplete story of mortgage applications.
<br> <br>

## Denial rate by race and ethnicity

This dashboard shows the total number of loans denied for the purchase
of a home as a primary residence. There are [multiple ways to determine
denial
rate](https://www.urban.org/sites/default/files/publication/33501/2000031-A-Better-Measure-of-Mortgage-Application-Denial-Rates.pdf),
but for simplicity we focus on the share of loans versus the total
number of loans for the same purpose.

Data is currently provided from 2018 to 2020 for all loans in Virginia
at the state, CBSA, and locality level. <br> <br>

```{r denials-race}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Denialratebyraceandethnicity/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br> <br>

## Additional resources

**Federal Reserve History**

[*Subprime Mortgage
Crisis*](https://www.federalreservehistory.org/essays/subprime-mortgage-crisis)

**Pew Research Center**

[*Blacks and Hispanics face extra challenges in getting home
loans*](https://www.pewresearch.org/fact-tank/2017/01/10/blacks-and-hispanics-face-extra-challenges-in-getting-home-loans/)

**Urban Institute**

[*Mortgage Denial Rates and Household Finances among Older
Americans*](https://www.urban.org/research/publication/mortgage-denial-rates-and-household-finances-among-older-americans)
<br> <br>

## Data sources

[Home Mortgage Disclosure
Act(HMDA)](https://www.consumerfinance.gov/data-research/hmda/)

Annual loan-level data releaed by the Consumer Financial Protection
Bureau (CFPB).

    ^^^ EVERYTHING ABOVE GOES ON PAGE ^^^
