---
title: "Sourcebook > Homeownership > Mortgages"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```

```
VVV EVERYTHING BELOW GOES ON PAGE VVV
```

## Home mortgages

The 1975 Home Mortgage Disclosure Act (HMDA) requires financial institutions to 
maintain and annually disclose data about home purchases and other loan activities. 
The primary purpose of this disclosure is to determine if financial institutions are 
serving the housing needs of communities and in a non-discriminatory manner. [HMDA data](https://www.consumerfinance.gov/data-research/hmda/) help provide an understanding 
of home purchase activity and allows us to look at home loans by race and ethnicity.

```
[add stock photo?]
```

### Total loans by race and ethnicity

This dashboard shows the total number of loans [originated](https://www.investopedia.com/terms/o/origination.asp)
for the purchase of a home as a primary residence. 

Data is currently provided from 2018 to 2020 for all loans in Virginia at the state, 
CBSA, and locality level.
<br>
<br>
```{r loans-race}
knitr::include_url(url = "https://public.tableau.com/app/profile/housingforward.virginia/viz/Sourcebook-Homeloansbyraceandethnicity/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>
```
[insert cautionary blurb about incomplete data]
```


### Denial rate by race and ethnicity

This dashboard shows the total number of loans denied for the purchase of a home 
as a primary residence. There are [multiple ways to determine denial rate](https://www.urban.org/sites/default/files/publication/33501/2000031-A-Better-Measure-of-Mortgage-Application-Denial-Rates.pdf), but for simplicity we focus on the share of loans versus the total number of loans
for the same purpose. 

Data is currently provided from 2018 to 2020 for all loans in Virginia at the state, 
CBSA, and locality level.
<br>
<br>
```{r denials-race}
knitr::include_url(url = "https://public.tableau.com/app/profile/housingforward.virginia/viz/Sourcebook-Denialratebyraceandethnicity/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>
## Data sources

[Home Mortgage Disclosure Act](https://www.consumerfinance.gov/data-research/hmda/)(HMDA)
Annual loan-level data releaed by the Consumer Financial Protection Bureau (CFPB). 


## Additional resources

```
[add relevant links]
```

```
^^^ EVERYTHING ABOVE GOES ON PAGE ^^^
```