---
title: "Sourcebook > Inventory > Housing type"
subtitle: "Data sources and methods"
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output:
  html_document:
    theme: lumen
    toc: true
    toc_float: true
    number_sections: true
---

# Data sources

### American Community Survey (ACS) {-}

The American Community Survey is an ongoing survey program of the U.S. Census Bureau. The ACS collects information about the nation’s social, economic, housing, and demographic characteristics on an annual basis. This information provides the most-up-to-date estimates between decennial censuses.

This section uses data from one ACS table:

* **Table B25032**: Tenure by Units in Structure

The most recent data used for this section are the **2015-2019 5-year estimates**. These estimates average all survey responses across this period to increase sample sizes and reduce margins of error.

Link: https://www.census.gov/programs-surveys/acs

# Setup

```{r setup, message=FALSE, warning=FALSE}

# Load packages and set global chunk options

library(tidycensus)
library(tidyverse)

knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)

```

# Data collection

```{r collect}

# Get variables for Table B25032

b25032_vars <- load_variables(2021, "acs5") %>% 
  filter(str_detect(name, "B25032")) %>% 
  filter(str_length(name) == 10)

# Get B25032 data for every locality in Virginia

b25032_raw <- get_acs(
  geography = "county",
  state = "VA",
  table = "B25032",
  year = 2021
)

```

# Data prep

```{r prep}

# Clean B25032 variable names

b25032_vars_clean <-  b25032_vars %>% 
  separate(label, into = c("est", "tot", "tenure", "type"),
           sep = "!!") %>% 
  select(variable = name, tenure, type) %>%
  filter(across(c(tenure, type), ~!is.na(.x))) %>% 
  mutate(across(.fns = ~str_remove_all(.x, ":")),
         tenure = str_remove_all(tenure, "-occupied housing units")) %>% 
  mutate(type = case_when(
    type == "1, detached" ~ "Single-family (detached)",
    type == "1, attached" ~ "Single-family (attached)",
    type == "2" ~ "Small-scale multifamily",
    type == "3 or 4" ~ "Small-scale multifamily",
    type == "5 to 9" ~ "Small-scale multifamily",
    type == "10 to 19" ~ "Small-scale multifamily",
    type == "20 to 49" ~ "Medium-scale multifamily",
    type == "50 or more" ~ "Large-scale multifamily",
    type == "Mobile home" ~ "Mobile (manufactured) home",
    type == "Boat, RV, van, etc." ~ "Other")
  )

# Join B25032 variables to data and calculate new sums

b25032_prep <- b25032_raw %>% 
  right_join(b25032_vars_clean, by = "variable") %>% 
  select(GEOID, tenure, type, estimate, moe) %>% 
  group_by(GEOID, tenure, type) %>% 
  summarise(
    estimate = sum(estimate),
    moe = moe_sum(moe, estimate)
  ) %>% 
  ungroup()

# Adorn subtotals for both owner and renter units and add data reliability info

b25032_data <- b25032_prep %>% 
  bind_rows(b25032_prep %>%  
    group_by(GEOID, type) %>% 
    summarise(
      estimate = sum(estimate),
      moe = moe_sum(moe, estimate)) %>% 
    ungroup()
  ) %>% 
  mutate(tenure = replace_na(tenure, "All")) %>% 
  mutate(cv = ((moe/1.645)/estimate)*100) %>% # Calculate coefficient of variation
  mutate(reliability = case_when(
    cv < 15 ~ "High",
    cv >= 15 & cv <= 30 ~ "Medium",
    cv > 30 ~ "Low")
  )

```

# Data export

```{r export}

# Save housing type data from B25032

write_csv(b25032_data, "data/housing-type.csv")

```

# Dashboard development

Reserved for future Shiny app development.