---
title: "Sourcebook > Economic > Poverty"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```

```
VVV EVERYTHING BELOW GOES ON PAGE VVV
```

## Poverty

The dashboards on this page include data on poverty rate by race and ethnicity, as
well as age. Poverty rate is the share of people whose income is below the poverty
line (i.e. the halfway point between no income and the median household income of 
the total population).

### Poverty rate by race and ethnicity

This dashboard shows poverty rate across Virginia from 2010 to 2020 by race and ethnicity. 


Data is available at the state, CBSA, and locality level.
<br>
<br>
```{r poverty-race}
knitr::include_url(url = "https://https://public.tableau.com/app/profile/housingforward.virginia/viz/Sourcebook-Povertybyraceandethnicity/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

### Poverty rate by age

This dashboard shows poverty rate across Virginia from 2010 to 2020 by age. It's important 
to note that the often high poverty rate among the 18 to 24 years old age group 

```{r poverty-age}
knitr::include_url(url = "https://public.tableau.com/app/profile/housingforward.virginia/viz/Sourcebook-Povertybyage/Statewide?:showVizHome=no&:embed=true", height = "500px")
```

## Data sources

[**American Community Survey**](https://www.census.gov/programs-surveys/acs) (ACS)

Period estimates based on the previous Decennial Census released annually. 


## Additional resources


```
[add relevant links]
```

```
^^^ EVERYTHING ABOVE GOES ON PAGE ^^^
```