---
title: "Sourcebook > Economic > Poverty"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```

```
VVV EVERYTHING BELOW GOES ON PAGE VVV
```
The dashboards on this page include data on poverty rate by race and ethnicity, as
well as age. Poverty rate is the share of people whose income is below the poverty
level as defined by the Census Bureau. 

## Poverty rate by race and ethnicity

This dashboard shows poverty rate across Virginia from 2010 to 2020 by race and ethnicity. 

Data is available at the state, CBSA, and locality level.
<br>
<br>
```{r poverty-race}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Povertybyraceandethnicity/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

## Poverty rate by age

This dashboard shows poverty rate across Virginia from 2010 to 2020 by age.

It is important to note that the poverty rate among the 18 to 24 age group is often much [higher than expected](https://www.census.gov/library/stories/2018/10/off-campus-college-students-poverty.html) in communities with a large number of off-campus college students, who have little to no reported income. While there are methods to exclude these individuals from a community's poverty rate (see "Additional resources" below), they are beyond the capacity for this dashboard to show.
<br>
<br>
```{r poverty-age}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Povertybyage/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

## Additional resources

**U.S. Census Bureau**

[*How the Census Bureau Measures Poverty*](https://www.census.gov/topics/income-poverty/poverty/guidance/poverty-measures.html)

**Brookings Institution**

[*Long shadows: The Black-white gap in multigenerational poverty*](https://www.brookings.edu/research/long-shadows-the-black-white-gap-in-multigenerational-poverty/)

**Fahe**

[*Appalachian Poverty*](https://fahe.org/appalachian-poverty/)

**National Poverty Center**

[*The Colors of Poverty: Why Racial & Ethnic Disparities Persist*](http://www.npc.umich.edu/publications/policy_briefs/brief16/PolicyBrief16.pdf)

**University of Virginia Weldon Cooper Center for Public Service**

[*How to modify poverty calculations for college towns*](https://statchatva.org/2016/03/07/how-to-modify-poverty-calculations-for-college-towns/)

## Data sources

[**American Community Survey**](https://www.census.gov/programs-surveys/acs) (ACS)

The American Community Survey is an ongoing survey program of the U.S. Census Bureau. The ACS collects information about the nation’s social, economic, housing, and demographic characteristics on an annual basis. This information provides the most-up-to-date estimates between decennial censuses.

This section uses data from one ACS table:

* **Table B17001**: Poverty Status by Sex by Age (iterations by race and ethnicity)

The most recent data used for this section are the **2016-2020 5-year estimates**. These estimates average all survey responses across this period to increase sample sizes and reduce margins of error. Data from the previous 5-year estimates are also used as a baseline to show change over time.

```
^^^ EVERYTHING ABOVE GOES ON PAGE ^^^
```