---
title: "Sourcebook > Affordability > Cost Burden"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```

```
VVV EVERYTHING BELOW GOES ON PAGE VVV
```

# Cost burden

While there are many different ways to measure housing affordability, the most common standard used by many planners and researches is **cost burden**. Cost burden is generally defined as paying more than 30 percent of household income for housing (rent or mortgage, plus the associated utilities). The dashboards on this page provide data on cost burden in Virginia between 2012 and 2018.

The total percentage of cost burden households consists of households that have no or negative income, are cost-burdened (spend between 30 and 50 percent of their income on housing), and severely cost-burdened (spend
more than 50 percent of their income on housing).
<br>
<br>

## Cost burden map

The following dashboard shows a map of cost burden across Virginia localities. You can filter
the map by year (2012 to 2018) and by tenure (homeowner or renter). Clicking on a 
locality or multiple localities will show you a detailed breakdown of cost-burdened
households by tenure.
<br>
<br>

```{r cb-map}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Housingcostburdenmap/cb_map_chart_db?:showVizHome=no&:embed=true", height = "500px")
```

<br>
<br>

## Cost burden by income

Lower-income households have less money to spend across the basic necessities we all need. Housing is the
number one expense among a majority of households, but households often have to make difficult choices
when it comes to choosing food on the table, getting adequate medical care, or getting to work.

The following dashboard shows cost burden by income group, defined as area median income (AMI) categories. Data is available at the state, CBSA, and locality level.
<br>
<br>

```{r cb-income}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Housingcostburdenbyincome/Statewide?:showVizHome=no&:embed=true", height = "500px")
```

<br>
<br>

## Cost burden by household type

Households spend their income differently depending on their household type. Seniors and families
with young children often need to spend more of their income on things like healthcare or childcare.
Because of these varying needs, different household types experience cost burden disproportionately.

The following dashboard shows cost burden by household type. Data is available at the state, CBSA, and locality level.
<br>
<br>

```{r cb-hhtype}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Housingcostburdenbyhouseholdtype/Statewide?:showVizHome=no&:embed=true", height = "500px")
```

<br>
<br>

## Cost burden by race

Systemic inequities have resulted in income and wealth disparities that have in turn left many households of color with disproportionate cost burden. 

The following dashboard shows cost burden by race and ethnicity. Data is available at the state, CBSA, and locality level.
<br>
<br>

```{r cb-race}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Housingcostburdenbyraceandethnicity/Statewide?:showVizHome=no&:embed=true", height = "500px")
```

<br>
<br>

## Additional resources

**Harvard Joint Center for Housing Studies**

[*Affordability research*](https://www.jchs.harvard.edu/research-areas/affordability)

**Greater Greater Washington**

[*What it means for a household to be cost-burdened*](https://ggwash.org/view/77638/housing-cost-burdens-explained_2)

## Data sources

[Comprehensive Housing Affordability Strategy](https://www.huduser.gov/portal/datasets/cp.html) (CHAS)

CHAS data are sourced from custom tabulations of the American Community Survey (ACS) 5-year estimates and are provided to the U.S. Department of Housing and Urban Development. Data are available at the locality level from 2009 to 2018.

Inconsistencies exist across data dictionaries before 2012. Therefore, Sourcebook only uses CHAS data from 2012 to 2018, when data variables are consistent.

```
^^^ EVERYTHING ABOVE GOES ON PAGE ^^^
```