---
title: "Sourcebook > Homeownership > House Price Index"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```


```
VVV EVERYTHING BELOW GOES ON PAGE VVV
```

# House Price Index

The dashboards on this show the House Price Index (HPI) for different markets of Virginia. The HPI is published by the Federal Housing Finance Agency (FHFA) and provides a standardized benchmark for single-family home values across a historical period. FHFA calculates the HPI using loan data from Freddie Mac and Fannie Mae.
<br>
<br>

## All-Transactions HPI

This dashboard shows the *All-Transactions House Price Index*, which is calculated using both:

* Prices recorded for home sale transactions, and
* Appraised values from home refinance mortgages.

Data is available by quarter for the state, Metropolitan Statistical Areas (MSA), and the nonmetropolitan areas of Virginia.

For statewide data, the HPI is indexed at 100 for the first quarter of 1980. For metro and nonmetro data, the HPI is indexed at 100 for the first quarter of 1995. This index is *not* seasonally adjusted.
<br>
<br>
```{r dashboard}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-HousePriceIndex/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

## Additional resources

**Federal Housing Finance Agency**

[*House Price Index Frequently Asked Questions*](https://www.fhfa.gov/Media/PublicAffairs/Pages/House-Price-Index-Frequently-Asked-Questions.aspx)

## Data sources

[**FHFA House Price Index**](link)

The House Price Index (HPI) is published by FHFA. This section uses the quarterly *All-Transactions Index* for the state, Metropolitan Statistical Areas (MSAs), and the nonmetropolitan area of Virginia. This index is estimated using both sales prices and appraisal data.


```
^^^ EVERYTHING ABOVE GOES ON PAGE ^^^
```