---
title: "Sourcebook > Homeownership > House Price Index"
subtitle: "Data sources and methods"
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output:
  html_document:
    theme: lumen
    toc: true
    toc_float: true
    number_sections: true
---

# Data sources

### House Price Index {-}

The House Price Index (HPI) is published by the Federal Housing Finance Agency (FHFA) and measures the average price change of existing single-family detached houses in repeat sales or refinancing of the same properties. This section uses the quarterly *All-Transactions Index* for the state, Metropolitan Statistical Areas (MSAs), and the nonmetropolitan area of Virginia. This index is estimated using both sales prices and appraisal data.

FHFA also publishes a "developmental" *HPI for Counties (All-Transactions Index)* dataset. Indexes this in data are annualized, and not available by quarter. Future Sourcebook updates may integrate this dataset.

Link: https://www.fhfa.gov/DataTools/Downloads/Pages/House-Price-Index-Datasets.aspx

# Setup

```{r setup, message=FALSE, warning=FALSE}

# Load packages and set global chunk options

library(tidyverse)
library(readxl)
library(httr)

knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

# Data collection

```{r collect}

# Read in the csv files for statewide and CBSA data

hpi_state_raw <- read_csv("https://www.fhfa.gov/DataTools/Downloads/Documents/HPI/HPI_AT_state.csv",
                   col_names = FALSE)

hpi_cbsa_raw <- read_csv("https://www.fhfa.gov/DataTools/Downloads/Documents/HPI/HPI_AT_metro.csv",
                     col_names = FALSE)

# Non-metro data is only available as an xls file; must read in via temp file

# Set up temp file container

temp <- tempfile(fileext = ".xls")

# Write nonmetro xls file to temp file on disk

httr::GET(url = "https://www.fhfa.gov/DataTools/Downloads/Documents/HPI/HPI_AT_nonmetro.xls",
          write_disk(temp))

# Import xls and skip header rows

hpi_nonmetro_raw <- read_excel(temp, sheet = 1, skip = 2)
```

# Data prep

```{r prep}

# Prep state data and filter only Virginia values

hpi_state_data <- hpi_state_raw %>% 
  setNames(c("name", "year", "quarter", "hpi")) %>% 
  filter(name == "VA") %>% 
  mutate(geography = "State",
         .before = 1) %>%
  mutate(fips = 51,
         .after = 2) %>% 
  mutate(name = "Virginia") %>% 
  add_column(stderror = NA)

# Prep CBSA data and filter only Virginia values

hpi_cbsa_data <- hpi_cbsa_raw %>% 
  setNames(c("name", "fips", "year", "quarter", "hpi", "stderror")) %>% 
  filter(str_detect(name, "VA")) %>% 
  mutate(geography = "CBSA",
         .before = 1) %>% 
  mutate(name = str_remove_all(name, "  \\(MSAD\\)")) %>% 
  mutate(stderror = str_replace_all(stderror, "[^0-9.]+", "")) %>% 
  mutate(hpi = na_if(hpi, "-"),
         stderror = na_if(stderror, "")) %>% 
  mutate(across(hpi:stderror, ~as.numeric(.x)))
  
# Prep non-metro data and filter only Virginia values

hpi_nonmetro_data <- hpi_nonmetro_raw %>% 
  setNames(c("name", "year", "quarter", "hpi", "stderror")) %>%
  filter(name == "VA") %>% 
  mutate(geography = "Nonmetro",
         .before = 1) %>%
  mutate(fips = NA,
         .after = 2) %>% 
  mutate(name = "Nonmetropolitan area")  %>% 
  mutate(stderror = str_replace_all(stderror, "[^0-9.]+", ""),
         stderror = as.numeric(stderror))

# Bind three geography levels together and create new date field with year and quarter

hpi_all_data <- bind_rows(hpi_state_data, hpi_cbsa_data, hpi_nonmetro_data) %>% 
  mutate(date = paste0(year, " Q", quarter),
         .after = 5)

```

# Data export

```{r export}

# Save HPI data

write_csv(hpi_all_data, "data/hpi.csv")

```

# Dashboard development

Reserved for future Shiny app development.