---
title: "Sourcebook > Rental > Assisted Rental"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```

```
VVV EVERYTHING BELOW GOES ON PAGE VVV
```

## Rental subsidy programs

The U.S. Department of Housing and Urban Development and U.S. Department of Agriculture run 
multiple federally subsidized housing programs throughout Virginia. These programs make up 
a bulk of the assisted rental housing in the state and are often layered over properties to 
maximize assistance. 

Thanks to the National Housing Preservation Database (NHPD) data from these two agencies and their
multiple programs are compiled in a single place. In addition, this data has been de-duplicated to 
provide a more accurate reflection of federally supported affordable housing properties.

The following map shows affordable housing properties across the Commonwealth and the type of 
subsidy used by those properties, as well as the total count of assisted units by subsidy type.

Clicking on a locality will change the breakdown of assisted units by type to reflect that locality. 
While, clicking on a subsidy type in the bottom frame will show where that subsidy is being utilized across
Virginia. 
<br>
<br>
```{r assisted}
knitr::include_url(url = "https://public.tableau.com/app/profile/housingforward.virginia/viz/Sourcebook-Assistedunits/AssistedProperties?:showVizHome=no&:embed=true", height = "800px")
```
<br>
<br>

It is important to note that NHPD uses data that has varying degrees of currency and 
accuracy. This is the result of inconsistency in data collection and different reporting 
schedules. NHPD estimates the number of assisted units, therefore the number of assisted units 
is an approximation.

The number of assisted units also does not include properties supported by Community Development 
Block Grant (CDBG), the National Housing Trust Fund (NHTF), McKinney-Vento Permanent Housing, HOPWA,
Tax-Exempt Multifamily Housing Bonds, or state programs like the Affordable and Special Needs Housing 
Program.

If you have more up-to-date information about a property, please contact eric@housingforwardva.org

## Additional resources

NHPD provides full program descriptions on their website [here](https://preservationdatabase.org/documentation/program-descriptions/).



## Data sources

[National Housing Preservation Database (NHPD)](https://preservationdatabase.org/)
De-duplicated data from HUD and USDA on affordable rental programs. Data are updated
tri-annually.

In addition, HousingForward Virginia does an additional data quality control measure and 
geocodes all addresses to ensure accuracy. This results in the removal of some properties 
due to our inability to verify a property. We also consult with Virginia Housing on inconclusive 
LIHTC properties.
