---
title: "Sourcebook > Rental > Assisted Rental"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```

```
VVV EVERYTHING BELOW GOES ON PAGE VVV
```
The U.S. Department of Housing and Urban Development (HUD) and U.S. Department of Agriculture (USDA) run 
multiple federally subsidized housing programs throughout Virginia. These programs make up most of the assisted rental housing in the state and are often combined with each other to maximize assistance for a single property.

Thanks to the National Housing Preservation Database (NHPD), data from these two agencies and their
multiple programs are compiled in a single place. In addition, this data has been de-duplicated to 
provide a more accurate reflection of federally supported affordable housing properties.

The following map shows affordable housing properties across the Commonwealth, the type of subsidy used by those properties, as well as the total count of assisted units by subsidy type.

Clicking on a locality will also filter the bottom chart. Clicking on a subsidy type will show where that subsidy is being used across Virginia. 
<br>
<br>
```{r assisted}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Assistedunits/AssistedProperties?:showVizHome=no&:embed=true", height = "800px")
```
<br>
<br>

It is important to note that NHPD data is built using data with varying degrees of currency and accuracy. This is the result of inconsistency in data collection and different reporting schedules across agencies. Numbers of assisted units by program are NHPD's best estimates.

NHPD's data collection also does not include properties supported by Community Development 
Block Grant (CDBG), the National Housing Trust Fund (NHTF), McKinney-Vento Permanent Housing, HOPWA,
tax-exempt multifamily housing bonds, certain state initiatives not using federal dollars, as well as local affordable dwelling unit ("inclusionary zoning") programs. However, affordable housing properties using these subsidies or incentives are likely to have also received support from the programs covered by NHPD.

## Additional resources

**NHPD documentation**

[*Program descriptions*](https://preservationdatabase.org/documentation/program-descriptions/)

[*Data dictionary*](https://preservationdatabase.org/documentation/data-dictionary/)

[*User guide*](https://preservationdatabase.org/documentation/user-guide/)

[*Data notes*](https://preservationdatabase.org/documentation/data-notes/)

[*Data sources*](https://preservationdatabase.org/documentation/data-sources/)

**NHPD reports**

[*National and state preservation profiles*](https://preservationdatabase.org/reports/preservation-profiles/)

[*Picture of Preservation report*](https://preservationdatabase.org/reports/picture-of-preservation/)

## Data sources

[National Housing Preservation Database](https://preservationdatabase.org/) (NHPD)

National database of affordable housing properties and their associated subsidies. To access the database and/or download data, users must register with a free account.

HousingForward conducts additional data quality control measures and geocodes all addresses to their associated city or county. This results in the omission of some properties we cannot fully verify. We also consult with Virginia Housing on LIHTC properties with subsidies marked "inconclusive."

Still, a full audit of each record is beyond our capacity. If you see any errors, or have more up-to-date information about a property, please contact [Eric Mai](mailto:eric@housingforwardva.org?subject=Property%20update%20for%20Sourcebook).

```
^^^ EVERYTHING ABOVE GOES ON PAGE ^^^
```