---
title: "Sourcebook > Rental > Assisted Rental"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```

```
VVV EVERYTHING BELOW GOES ON PAGE VVV
```

## Rental subsidy programs

The U.S. Department of Housing and Urban Development and U.S. Department of Agriculture run 
multiple federally subsidized housing programs throughout Virginia. These programs make up 
a bulk of the assisted rental housing in the state and are often layered over properties to 
maximize assistance. 

Thanks to the National Housing Preservation Database (NHPD) data from these two agencies and their
multiple programs are compiled in a single place. In addition, this data has been de-duplicated to 
provide a more accurate reflection of federally supported affordable housing properties.

The following map shows affordable housing properties across the Commonwealth and the type of 
subsidy used by those properties, as well as the total count of assisted units by subsidy type.

Clicking on a locality will change the breakdown of assisted units by type to reflect that locality. 
While, clicking on a subsidy type in the bottom frame will show where that subsidy is being utilized across
Virginia. 
<br>
<br>
```{r assisted}
knitr::include_url(url = "https://public.tableau.com/app/profile/housingforward.virginia/viz/Sourcebook-Assistedunits/AssistedProperties?:showVizHome=no&:embed=true", height = "800px")
```
<br>
<br>

## Data sources

[National Housing Preservation Database (NHPD)](https://preservationdatabase.org/)
De-duplicated data from HUD and USDA on affordable rental programs. Data are updated
tri-annually.

In addition, HousingForward Virginia does an additional data quality control measure and 
geocodes all addresses to ensure accuracy. This results in the removal of some properties 
due to our inability to verify a property.
