---
title: "Sourcebook > Inventory > Housing characteristics"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```


```
VVV EVERYTHING BELOW GOES ON PAGE VVV
```

## Housing characteristics

The dashboards on this page break down housing units by selected physical characteristics. These data are helpful in understanding the design and size of a community's housing stock, and whether that supply adequately meeds the needs of the population.
<br>
<br>
```
[add stock photo?]
```

### Number of bedrooms by tenure

This dashboard shows the current supply of occupied housing units among owners, renters, and all households, broken down by the number of bedrooms each home contains.

Data is available at the state, CBSA, and locality level.
<br>
<br>
```{r dashboard}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Numberofbedroomsbytenure/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

## Data sources

[**American Community Survey**](https://www.census.gov/programs-surveys/acs)

The American Community Survey is an ongoing survey program of the U.S. Census Bureau. The ACS collects information about the nation’s social, economic, housing, and demographic characteristics on an annual basis. This information provides the most-up-to-date estimates between decennial censuses.

This section uses data from one ACS table:

* **Table B25042**: Tenure by Bedrooms

The most recent data used for this section are the **2015-2019 5-year estimates**. These estimates average all survey responses across this period to increase sample sizes and reduce margins of error.

## Additional resources

```
[add relevant links]
```

```
^^^ EVERYTHING ABOVE GOES ON PAGE ^^^
```