---
title: "Sourcebook > Affordability > HUD AMI limits"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```


```
VVV EVERYTHING BELOW GOES ON PAGE VVV
```

## Area Median Income limits

HUD calculates income limits to determine whether households are eligible for federally-funded housing assistance programs, including public housing, Section 8 project-based, Housing Choice Vouchers, and others. These limits are created using American Community Survey data on household incomes from the Census Bureau.

Using the median family income as the "Area Median Income" (AMI), HUD produces income limits for 30, 50, and 80 percent of AMI, further broken down by number of persons within a household. These levels correspond to common terms to describe these income ranges:

* Between 80 and 50 percent AMI is *low-income*,
* Between 50 and 30 percent AMI is *very low-income*, and
* Below 30 percent AMI is *extremely low-income*.

Income limits for localities in metropolitan areas are provided at that regional level, so a central city and adjacent suburban county will often share AMI values. Non-metropolitan localities have their AMI limits calculated separately.
<br>
<br>

### HUD AMI limits

This dashboard shows HUD AMI limits by income category and household size for every locality in Virginia. If a locality is in an area where AMI values are set regionally, the chart title will display that geography's name. Prior year AMI limits are shown back to 2017.
<br>
<br>
```{r dashboard}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-HUDAMIlimits/HUD-AMI?:language=en-US&:display_count=n&:origin=viz_share_link", height = "500px")
```
<br>
<br>

## Data sources

[**U.S. Department of Housing and Urban Development**](https://www.huduser.gov/portal/datasets/il.html)

HUD publishes new income limits in April of each year.

## Additional resources

[**Multifamily Tax Subsidy Projects (MTSP) Income Limits**](https://www.huduser.gov/portal/datasets/mtsp.html)

Additional income limits published by HUD as guidance for LIHTC program and other projects where rent rates are set using AMI limits at additional income levels, such as 60 percent of AMI.

```
^^^ EVERYTHING ABOVE GOES ON PAGE ^^^
```