---
title: "homeownership rate"
author: "Eric Mai"
date: "1/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Sources
Homeownership rate is derived by dividing the total number of owner-occupied
housing units by the total number of occupied housing units (owner-occupied +
renter-occupied). The source of this data set is the Table B25003 from the 
American Community Survey.

Table B25003 is also provided by race and ethnicity by applying letters A 
through I to the end of the table ID. 

+ **A** is White Alone
+ **B** is Black or African American Alone
+ **C** is American Indian and Alaska Native Alone
+ **D** is Asian Alone
+ **E** is Native Hawaiian and Other Pacific Islander Alone
+ **F** is Some Other Race Alone
+ **G** is Two or More Races
+ **H** is White Alone, Not Hispanic or Latino
+ **I** is Hispanic or Latino

## Data Collection and Data Prep
The following script pulls Table B25003 for all races and ethnicity, and cleans
the data in a single script.

``````{r Download B25003, echo=T, message=FALSE, warning=FALSE, paged.print=FALSE, results='hide'}

#Load the the necessary packages to download the data from Census API.
library(tidyverse)
library(tidycensus)

# Create an object for all years.
years <- 2010:2019

# Create an object for all tables needed. This includes a table that aggregates 
# to total housing units.
b25003 <- c("B25003", paste0("B25003", LETTERS[2:9]))

# Create a function to convert variable to race or ethnicity variable.
concept_to_race <- function(x) {
  out <- x %>%
    str_remove_all("HOUSEHOLDER") %>%
    str_remove_all("TENURE \\(|\\)") %>%
    str_to_title() %>%
    str_replace_all("And", "and") %>%
    str_replace_all("Or", "or") %>%
    str_remove_all(" Alone")
  
  out
}

# Pull the table variables, excluding Puerto Rico.
b25003_defns <- load_variables(2019, "acs5") %>%
  filter(str_sub(name, end = 6) %in% b25003) %>% 
  filter(str_detect(name, "PR") == FALSE)

# Clean the variables provided by Census API and separate needed variables into
# appropriate columns.
b25003_cleaned <- b25003_defns %>%
  mutate(race = concept_to_race(concept)) %>%
  separate(label, into = c("est", "total", "tenure"), sep = "!!") %>% 
  select(variable = name, race, tenure) %>% 
  mutate(across(.fns = ~replace_na(.x, "All")),
         across(.fns = ~str_remove_all(.x, ":")),
         across(.fns = ~str_remove_all(.x, " --")),
         across(.fns = ~str_replace_all(.x, "total", "All")),
         across(.fns = ~str_replace_all(.x, "Tenure", "All")))

# Download the data for all counties in Virginia. Functions interate across all
# tables and all years. If desiring to change to state then input "state" instead
# of "county" and remove 'state = "VA" from function.
output_b25003 <- map_dfr(b25003, function(tb) {
  yearly_data <- map_dfr(years, function(yr) {
    
    acs_pull <- get_acs(
      geography = "county",
      table = tb,
      year = yr,
      state = "VA"
    ) %>%
      left_join(b25003_cleaned, by = "variable")
    
    acs_rearranged <- acs_pull %>%
      mutate(year = yr) %>%
      select(variable, year, locality = NAME, fips = GEOID, race, tenure,
             estimate, moe)
    
    acs_rearranged
  })
  yearly_data
})

# The previous data output makes it difficult to easily calculate dynamic 
# homeownership rates if wanting to combine multiple localities or races. In
# order to create these calculations more easily, unpivot the columns to rows
# utilizing the pivote_wider function.
output_b25003_wide <- output_b25003 %>% 
  select( -c(variable)) %>% 
  pivot_wider(names_from = tenure,
              values_from = c(estimate, moe)) %>% 
  select(year:race, est_all = estimate_All, est_owner = "estimate_Owner occupied",
         est_renter = "estimate_Renter occupied", moe_all = moe_All, 
         moe_owner = "moe_Owner occupied",
         moe_renter = "moe_Renter occupied")

# Clean the locality column to remove comma and Virginia from data, as well as
# converting Bedford city (51515) to Bedford County (51019).

output_b25003_wide <- output_b25003_wide %>%
  mutate(across(.fns = ~str_remove_all(.x, ", Virginia")),
         across(.fns = ~str_replace_all(.x, "Bedford city", "Bedford County")),
         across(.fns = ~str_replace_all(.x, "51515", "51019")))


# Write to csv.
write_csv(output_b25003_wide, "b25003_wide.csv")

```

