---
title: "Sourcebook > Economic > Household income"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```

    VVV EVERYTHING BELOW GOES ON PAGE VVV

## Household income

The dashboards on this page show the changing incomes of Virginia residents by tenure, age, race, and ethnicity. This information helps inform advocates and practitioners on the changing economic status of communities and the growing need to address both housing and economic development.

```
[add stock photo?]
```

```
*[Insert info box on using medians]*
Using medians instead of the mean, or average, is a much better statistic to use because it helps accounts for data that has outliers. This is important for household income because a few really wealth individuals living in a county could raise a mean household income higher --- making it seem like more people in the county are wealthier when they are not. 

Instead, the median is better at capturing the "typical" household because it is the the middle value --- meaning that half of households are below that value and the other half are above. This removes the influence that outliers.

```

### Median household income by tenure

This dashboard shows the median household income of homeowners, renters, and both tenures combined from 2010 to 2019. Dollar figures have been adjusted to the most recent inflation-adjusted value which is available via the American Community Survey. 

Data is available at the state, CBSA, and locality level.
<br>
<br>
```{r med-hh-income-tenure}
knitr::include_url(url = "https://public.tableau.com/app/profile/housingforward.virginia/viz/Sourcebook-Medianhouseholdincomebytenure/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

### Income distribution by tenure

This dashboard shows the distribution of households by income group and tenure. Dollar figures have not been adjusted because the focus of the data is the general distribution of households, not comparing a dollar value across time.

Data is available at the state, CBSA, and locality level.
<br>
<br>
```{r income-tenure}
knitr::include_url(url = "https://public.tableau.com/app/profile/housingforward.virginia/viz/Sourcebook-Incomedistributionbytenure/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

### Median household income by age

This dashboard shows the median household income of by age of householder from 2010 to 2019. Dollar figures have been adjusted to the most recent inflation-adjusted value which is available via the American Community Survey. 

Data is available at the state, CBSA, and locality level.
<br>
<br>
```{r med-hh-income-age}
knitr::include_url(url = "https://public.tableau.com/app/profile/housingforward.virginia/viz/Sourcebook-Medianhouseholdincomebyage/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

### Median household income by race and ethnicity

This dashboard shows the median household income of by race and ethnicity of householder from 2010 to 2019. Dollar figures have been adjusted to the most recent inflation-adjusted value which is available via the American Community Survey. 

All race categories presented are non-Hispanic, while the Hispanic or Latino category represents any race.

Data is available at the state, CBSA, and locality level.

<br>
<br>
```{r med-hh-income-race}
knitr::include_url(url = "https://public.tableau.com/app/profile/housingforward.virginia/viz/Sourcebook-Medianhouseholdincomebyraceandethnicity/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>


## Data sources

[**American Community Survey**](https://www.census.gov/programs-surveys/acs) (ACS)

Period estimates based on the previous Decennial Census released annually. 


## Additional resources

```
[add relevant links]
```

```
^^^ EVERYTHING ABOVE GOES ON PAGE ^^^
```