---
title: "Sourcebook > Economic > Household income"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```

    VVV EVERYTHING BELOW GOES ON PAGE VVV

# Household income

The dashboards on this page show the changing incomes of Virginia residents by tenure, age, race, and ethnicity. This information helps inform advocates and practitioners on the changing economic status of communities and the growing need to address both housing and economic development.

```
[add stock photo?]
```


For most socioeconomic data, medians are preferred over means to show the average value for a given sample. Because medians represent the literal "middle" value, they are less influenced by outliers than means. This is important for household income, because a small number of very wealthy individuals could inflate the mean household income for a given community. 

Interpreting medians is straightforward. For example, a county with a median household income of $75,000 has one half of all households earning below that amount, while the other half earns more.
<br>
<br>

## Median household income by tenure

This dashboard shows the median household income of homeowners, renters, and both tenures combined from 2010 to 2020. Dollar figures have been adjusted to the most recent inflation-adjusted value which is available via the American Community Survey. 

Data is available at the state, CBSA, and locality level.
<br>
<br>
```{r med-hh-income-tenure}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Medianhouseholdincomebytenure/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

## Income distribution by tenure

This dashboard shows the distribution of households by income group and tenure. Dollar figures have not been adjusted because the focus of the data is the general distribution of households, not comparing a dollar value across time.

Data is available at the state, CBSA, and locality level.
<br>
<br>
```{r income-tenure}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Incomedistributionbytenure/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

## Median household income by age

This dashboard shows the median household income of by age of householder from 2010 to 2020. Dollar figures have been adjusted to the most recent inflation-adjusted value which is available via the American Community Survey. 

Data is available at the state, CBSA, and locality level.
<br>
<br>
```{r med-hh-income-age}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Medianhouseholdincomebyage/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

## Median household income by race and ethnicity

This dashboard shows the median household income of by race and ethnicity of householder from 2010 to 2020. Dollar figures have been adjusted to the most recent inflation-adjusted value which is available via the American Community Survey. 

All race categories presented are non-Hispanic, while the Hispanic or Latino category represents any race.

Data is available at the state, CBSA, and locality level.

<br>
<br>
```{r med-hh-income-race}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Medianhouseholdincomebyraceandethnicity/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

## Additional resources

**Brookings Institution**

[*Examining the Black-white wealth gap*](https://www.brookings.edu/blog/up-front/2020/02/27/examining-the-black-white-wealth-gap/)

**National Public Radio (NPR)**

[*There Is Growing Segregation In Millennial Wealth*](https://www.npr.org/sections/money/2021/04/27/990770599/there-is-growing-segregation-in-millennial-wealth)

**U.S. Census Bureau**

[*Many U.S. Households Do Not Have Biggest Contributors to Wealth: Home Equity and Retirement Accounts*](https://www.census.gov/library/stories/2019/08/gaps-in-wealth-americans-by-household-type.html?utm_campaign=20190827msacos1ccstors&utm_medium=email&utm_source=)

## Data sources

[**American Community Survey**](https://www.census.gov/programs-surveys/acs) (ACS)

The American Community Survey is an ongoing survey program of the U.S. Census Bureau. The ACS collects information about the nation’s social, economic, housing, and demographic characteristics on an annual basis. This information provides the most-up-to-date estimates between decennial censuses.

This section uses data from four ACS tables:

* **Table B25119**: Median Household Income the Past 12 Months by Tenure
* **Table B25010**: Tenure by Household Income
* **Table B19049**: Median Household Income by Age of Householder
* **Table B19031**: Median Household Income (iterations by race and ethnicity)

The most recent data used for this section are the **2016-2020 5-year estimates**. These estimates average all survey responses across this period to increase sample sizes and reduce margins of error. Data from the previous 5-year estimates are also used as a baseline to show change over time.



```
^^^ EVERYTHING ABOVE GOES ON PAGE ^^^
```