---
title: "Sourcebook > Inventory > Overcrowding"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```


```
VVV EVERYTHING BELOW GOES ON PAGE VVV
```

## Overcrowding

The dashboards on this page break down households by the number of occupants per room. These data are helpful in understanding how many households are living in homes that are too small for them to live in comfortably. People living in overcrowded conditions are more likely to experience physical and mental health issues---including an increased likelihood of disease transmission.

While definitions vary, the common “overcrowded” threshold is 1.0 persons per room and the “very overcrowded” threshold is 1.5 persons per room.
<br>
<br>
```
[add stock photo?]
```

### Overcrowded households by tenure

This dashboard shows the share of households that are overcrowded or very overcrowded among owners, renters, and all households. 

Data is available at the state, CBSA, and locality level.
<br>
<br>
```{r dashboard}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Overcrowdedhouseholdsbytenure/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

## Additional resources

**U.S. Department of Housing and Urban Development**

[*Measuring Overcrowding in Housing*](https://www.census.gov/programs-surveys/ahs/research/publications/Measuring_Overcrowding_in_Hsg.html) (PDF)

**Urban Institute**

[*America’s Housing Is Getting More Crowded. How Will That Affect Children?*](https://www.urban.org/urban-wire/americas-housing-getting-more-crowded-how-will-affect-children)

**National Multifamily Housing Council**

[*How Much is Overcrowding a Factor in COVID-19 Risk*](https://www.nmhc.org/research-insight/research-notes/2020/how-much-is-overcrowding-a-factor-in-covid-19-risk/)

## Data sources

[**American Community Survey**](https://www.census.gov/programs-surveys/acs)

The American Community Survey is an ongoing survey program of the U.S. Census Bureau. The ACS collects information about the nation’s social, economic, housing, and demographic characteristics on an annual basis. This information provides the most-up-to-date estimates between decennial censuses.

This section uses data from one ACS table:

* **Table B25014**: Tenure by Occupants Per Room

The most recent data used for this section are the **2015-2019 5-year estimates**. These estimates average all survey responses across this period to increase sample sizes and reduce margins of error.

```
^^^ EVERYTHING ABOVE GOES ON PAGE ^^^
```