---
title: "Sourcebook > Demographics > Household size"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```

```
VVV EVERYTHING BELOW GOES ON PAGE VVV
```

## Household size

The dashboards on this page break down the households by the number of people in them, including all adults and children. These data are helpful in understanding the different types, sizes, and locations of homes that a community's population may need.
<br>
<br>
```
[add stock photo?]
```

### Households by size

This dashboard shows households by size in 2010 and 2019. The label on the 2019 values show the percent change in the number of households of that size from 2010.

Data is available at the state, CBSA, and locality level. Use the filter to view all households, owner-occupied households, or renter-occupied households.
<br>
<br>
```{r pop-total}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Householdsbysize/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

### Average household size

This dashboard shows the average household size for all households, owner-occupied households, and renter-occupied households in 2010 and 2019. This value is calculated by dividing the total population in occupied housing units by the number of households.

Data is available at the state, CBSA, and locality level.
<br>
<br>
```{r pop-components}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Averagehouseholdsize/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

## Additional resources

**Pew Research Group**

[*The number of poeple in the average U.S. household is going up for the first time in over 160 years*](https://www.pewresearch.org/fact-tank/2019/10/01/the-number-of-people-in-the-average-u-s-household-is-going-up-for-the-first-time-in-over-160-years/)

**Population Reference Bureau (PRB)**

[*U.S. household composition shifts as population grows older; more young adults with parents*](https://www.prb.org/resources/u-s-household-composition-shifts-as-the-population-grows-older-more-young-adults-live-with-parents/)


## Data sources

[**American Community Survey**](https://www.census.gov/programs-surveys/acs) (ACS)  

The American Community Survey is an ongoing survey program of the U.S. Census Bureau. The ACS collects information about the nation’s social, economic, housing, and demographic characteristics on an annual basis. This information provides the most-up-to-date estimates between decennial censuses.

This section uses data from two ACS tables:

* **Table B25009**: Tenure by Household Size
* **Table B25010**: Average Household Size of Occupied Housing Units by Tenure

The most recent data used for this section are the **2015-2019 5-year estimates**. These estimates average all survey responses across this period to increase sample sizes and reduce margins of error. Data from the **2006-2010 5-year estimates** are also used as a baseline to show change over time.


```
^^^ EVERYTHING ABOVE GOES ON PAGE ^^^
```