---
title: "Sourcebook > Demographics > Household size"
subtitle: "Data sources and methods"
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output:
  html_document:
    theme: lumen
    toc: true
    toc_float: true
    number_sections: true
---

# Data sources

### American Community Survey (ACS) {-}

The American Community Survey is an ongoing survey program of the U.S. Census Bureau. The ACS collects information about the nation’s social, economic, housing, and demographic characteristics on an annual basis. This information provides the most-up-to-date estimates between decennial censuses.

This section uses data from two ACS tables:

* **Table B25009**: Tenure by Household Size
* **Table B25010**: Average Household Size of Occupied Housing Units by Tenure

The most recent data used for this section are the **2015-2019 5-year estimates**. These estimates average all survey responses across this period to increase sample sizes and reduce margins of error. Data from the **2006-2010 5-year estimates** are also used as a baseline to show change over time.

Link: https://www.census.gov/programs-surveys/acs

# Setup

```{r setup, message=FALSE, warning=FALSE}

# Load packages and set global chunk options

library(tidycensus)
library(tidyverse)

knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

# Data collection

```{r collect}

# Set year range for data collection

years <- c(2010, 2022)

# Get variables for Table B25009

b25009_vars <- load_variables(2022, "acs5") %>% 
  filter(str_detect(name, "B25009"))

# Get B25009 data for every locality in Virginia from 2010 and 2019

b25009_raw <- map_dfr(years, function(yr) {
  b25009_pull <- get_acs(
    geography = "county",
    state = "VA",
    table = "B25009",
    year = yr) %>% 
    mutate(year = yr)
})

# Get variables for Table B25010

b25010_vars <- load_variables(2022, "acs5") %>% 
  filter(str_detect(name, "B25010"))

# Average household size data from B25010 cannot be aggregated to CBSA or state level
# Must pull B25010 for each geographic level separately
# Create list of CBSA codes for all that are in or partly in Virginia

cbsa <- c("13720", "13980", "14140", "16820", "19260", "25500", "28700", "31340", "32300", "40060", "40220", "44420", "47260", "47900", "49020")

# Get B25010 data and add 'geography' column

b25010_raw <- map_dfr(years, function(yr) {
  
  # Pull data for state
  
  b25010_state <- get_acs(
    geography = "state",
    state = "VA",
    table = "B25010",
    year = yr) %>% 
    mutate(year = yr,
           geography = "state")
  
  # Pull data for CBSAs
  
  b25010_cbsa <- get_acs(
    geography = "cbsa",
    table = "B25010",
    year = yr) %>% 
    mutate(year = yr,
           geography = "cbsa") %>% 
    filter(GEOID %in% cbsa)
  
  # Pull data for localities
  
  b25010_locality <- get_acs(
    geography = "county",
    state = "VA",
    table = "B25010",
    year = yr) %>% 
    mutate(year = yr,
           geography = "locality")
  
  # Bind dataframes together
  
  b25010_all <- bind_rows(b25010_state, b25010_cbsa, b25010_locality)
  
})

```

# Data prep

```{r prep}

# Clean B25009 variable names

b25009_vars_clean <- b25009_vars %>% 
  separate(label, into = c("est", "tot", "tenure", "hhsize"),
           sep = "!!") %>% 
  select(variable = name, tenure, hhsize) %>% 
  mutate(across(.fns = ~str_remove_all(.x, ":")),
         tenure = str_remove_all(tenure, " occupied"),
         hhsize = str_remove_all(hhsize, " household")) %>%
  filter(across(c(tenure, hhsize), ~ !is.na(.x))) %>% 
  mutate(hhsize = case_when(
    str_detect(hhsize, "[4-7]") ~ "4-or-more person",
    TRUE ~ hhsize)
    )

# Join B25009 variables to data, calculate new sums, and add data reliability info

b25009_data <- b25009_raw %>%
  mutate(GEOID = str_replace_all(GEOID, "51515", "51019")) %>% # Merge 'Bedford city' values from 2010 to 'Bedford county'
  right_join(b25009_vars_clean, by = "variable") %>% # Use right_join to leave out unnecessary subtotals from data
  select(GEOID, year, tenure, hhsize, estimate, moe) %>% # Simplify columns
  group_by(GEOID, year, tenure, hhsize) %>% # Collapse and sum hhsizes
  summarise(
    estimate = sum(estimate),
    moe = moe_sum(moe, estimate) # Calculate new group-wise margins of error 
  ) %>% 
  ungroup() %>% 
  mutate(cv = ((moe/1.645)/estimate)*100) %>% # Calculate coefficient of variation
  mutate(reliability = case_when(
    cv < 15 ~ "High",
    cv >= 15 & cv <= 30 ~ "Medium",
    cv > 30 ~ "Low")
  )

# Clean B25010 variable names

b25010_vars_clean <- b25010_vars %>% 
  separate(label, into = c("est", "avg", "tot", "tenure"),
           sep = "!!") %>% 
  select(variable = name, tenure) %>% 
  mutate(across(.fns = ~str_remove_all(.x, ":")),
         across(.fns = ~replace_na(.x, "All")),
         tenure = str_remove_all(tenure, " occupied"))

# Join B25010 variables to data, calculate new sums, and add data reliability info

b25010_data <- b25010_raw %>%
  filter(!GEOID == "51515") %>% # Filter out 2010 'Bedford city' values since 2019 values do not exist
  right_join(b25010_vars_clean, by = "variable") %>% # Join variables to data
  select(GEOID, geography, year, tenure, estimate, moe) %>% # Simplify columns
  mutate(cv = ((moe/1.645)/estimate)*100) %>% # Calculate coefficient of variation
  mutate(reliability = case_when(
    cv < 15 ~ "High",
    cv >= 15 & cv <= 30 ~ "Medium",
    cv > 30 ~ "Low")
  )

```

# Data export

```{r export}

# Save household size data from B25009

write_csv(b25009_data, "data/hh-size.csv")

# Save average household size data from B25010

write_csv(b25010_data, "data/hh-size-avg.csv")

```

# Dashboard development

Reserved for future Shiny app development.