---
title: "Sourcebook > Economic > Employment"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```

```
VVV EVERYTHING BELOW GOES ON PAGE VVV
```
# Employment and housing

The dashboards on this page include data on employment levels, unemployment rate, 
and the jobs-housing balance across Virginia from 2008 to 2020. Jobs and 
housing are deeply intertwined. Without an adequate housing supply, communities 
can struggle to attract jobs and economic growth. This can lead to increases in 
commute times, greater traffic impacts, and sprawling development. 


```
[add stock photo?]
```

## Employment levels

The following dashboard shows the employment levels across Virginia from 2008 to 2020
by month. The data have not been seasonally adjusted, therefore you will see spikes 
across time that reflect the influences of seasonal patterns in employment and 
unemployment. 

```{r employment}
knitr::include_url(url = "https://public.tableau.com/app/profile/housingforward.virginia/viz/Sourcebook-VirginiaEmploymentLevels/State?:showVizHome=no&:embed=true", height = "500px")
```


## Unemployment rate

The following dashboard shows the unemployment rate from 2008 through 2021. Unemployment
rate is the share of people in the labor force (all people age 16 and older who are 
working or looking for work) that do not have a job but are actively looking for work. 

```{r unemployment}
knitr::include_url(url = "https://public.tableau.com/app/profile/housingforward.virginia/viz/Sourcebook-VirginiaUnemploymentRate/Statewide?:showVizHome=no&:embed=true", height = "500px")
```

## Jobs-housing growth

When the number of jobs are growing faster than the supply of housing, communities can face 
challenges in housing affordability. Growing demand from new employees can drive housing 
demand up and impact existing residents. But slow housing growth can also have a negative impact on economic growth within communities. 

```{r jobs-housing}
knitr::include_url(url = "https://public.tableau.com/app/profile/housingforward.virginia/viz/Sourcebook-Jobs-housinggrowth/Statewide?:showVizHome=no&:embed=true", height = "500px")
```

## Data sources

[Local Area Unemployment Statistics](https://www.bls.gov/lau/) (LAUS)

Monthly and annual estimates released by the Bureal of Labor Statistics. Available at the locality level from
1976 to 2021.


[Census Population Estimates Program](https://www.census.gov/programs-surveys/popest.html) (PEP)

Annual estimates of total population released by the U.S. Census Bureau. Available at locality level from 2010 to 2019. Estimates by age, race, and ethnicity also available, along with components of population change.

[2020 Decennial Census](https://www.census.gov/programs-surveys/decennial-census/about/rdo/summary-files.html)

Initial data from the 2020 Decennial Census is available in the PL 94-171 Redistricting summary file. As of February 2022, we use this data to append 2020 population counts onto the PEP time series data through 2019. Complete 2010 to 2020 Intercensal Estimates will be published in late 2022.

## Additional resources

[Jobs-Housing Balance](http://planning-org-uploaded-media.s3.amazonaws.com/publication/download_pdf/PAS-Report-516.pdf)


```
^^^ EVERYTHING ABOVE GOES ON PAGE ^^^
```
