---
title: "Sourcebook > Homeownership > Home sales"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```


```
VVV EVERYTHING BELOW GOES ON PAGE VVV
```

# Home sales

The dashboards on this page track the for-sale component of Virginia's housing market. These data are helpful in understanding where homeownership opportunities might be within or beyond the reach for today's homebuyers.

Sales data on this page is presented by quarter, beginning with 2014 Q1. New quarterly data will be uploaded approximately two months after the end of each quarter.

Home sales in this data capture detached single-family homes, attached single-family homes (townhomes), and condo units.
<br>
<br>

## Median sales price

This dashboard shows the median sales price for all homes sold by quarter.

Data is available at the state, MSA, and locality level.
<br>
<br>
```{r msp}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Mediansalesprice/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

## Median days on market

This dashboard shows the median days on market for all homes sold by quarter. Lower values are indicative of a tighter market as buyers place offers and close on homes quickly, often competing with others. Markets with higher values generally have lower demand, as homes take longer times to sell.

Data is available at the state, MSA, and locality level.
<br>
<br>
```{r mdom}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Mediandaysonmarket/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

## Additional resources

[**Virginia REALTORS速 Home Sales Reports**](https://virginiarealtors.org/research/reports/home-sales-reports/)

Monthly statewide and regional reports on the homeownership market in Virginia, provided with commentary and analysis from researchers at Virginia REALTORS速.

## Data sources

[**Virginia REALTORS速**](https://virginiarealtors.org/research/data/)

Virginia REALTORS速 and [R E Stats Inc.](https://www.redatum.com/) provide HousingForward Virginia with quarterly data collected from multiple listing services (MLS) across the state.


```
^^^ EVERYTHING ABOVE GOES ON PAGE ^^^
```