---
title: "Sourcebook > Inventory > Production"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```


```
VVV EVERYTHING BELOW GOES ON PAGE VVV
```
This page includes dashboards that show housing production trends across Virginia using annual residential building permits.
<br>
<br>

## Annual building permits by structure type

This dashboard shows the annual number of residential building permits issued by local governments in Virginia from 2000 to 2020.

Use the filter to select between some or all the following building types:

* **1-unit** buildings (single-family homes)
* **2-unit** buildings (duplexes)
* **3-4 unit** buildings (triplexes and quadraplexes)
* **5+ unit** buildings (larger apartment buildings)

Data is available at the state, CBSA, and locality level.
<br>
<br>
```{r dashboard}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Annualbuildingpermitsbystructuretype/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

## Additional resources

**Business Insider**

[*Housing in the US has not been able to keep up with buyer demand over the past decade*](https://www.businessinsider.com/us-underbuilding-housing-over-the-past-decade-2020-9)

[**Home Builders Association of Virginia**](https://www.hbav.com/)

[**National Association of Home Builders**](https://www.nahb.org/)

## Data sources

[**Census Building Permits Survey**](https://www.census.gov/construction/bps/)

Total number of building permits issued by localities, collected and released by the U.S. Census Bureau. Data is available annually, year-to-date, and monthly. Available geographies include states, selected metropolitan areas, localities and places.

```
^^^ EVERYTHING ABOVE GOES ON PAGE ^^^
```