---
title: "Sourcebook > Rental > Rental vacancy"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```

```
VVV EVERYTHING BELOW GOES ON PAGE VVV
```
The dashboards on this page show the rental vacancy across Virginia from 2010 to 2020. Rental vacancy rates are a common indicator of rental demand in a community. 

Although there are [different measures](https://www.stessa.com/blog/vacancy-rate/) of rental vacancy,
a general measure is the share of vacant rental units versus the total share of rental units in an area (both vacant and occupied). A low rental vacancy rate can indicate strong demand for rental housing. Limited supply will often exacerbate affordability challenges, while a high rental vacancy rate can indicate a weak rental 
market.

## Rental vacancy rate

The following dashboards shows the rental vacancy rate based on American Community Survey data for all Virginia localities from 2010 to 2020. This measure focuses on the physical vacancy rate. This is the share of vacant rental units (units that are not currently occupied) versus the total number of rental units (both vacant and occupied).

Data is available at the state, CBSA, and locality level.
<br>
<br>
```{r acs-vacancy}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-RentalvacancyrateACS/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

## Additional resources

**Investopedia**

[*Vacancy Rate Definition*](https://www.investopedia.com/terms/v/vacancy-rate.asp)

**Strong Towns**

[*What Vacancy Rates Tell You About a Housing Shortage (And What They Don't)*](https://www.strongtowns.org/journal/2020/8/30/what-vacancy-rates-tell-you-about-a-housing-shortage)

**U.S. Census Bureau**

[*Housing Vacancies and Homeownership (CPS/HVS)*](https://www.census.gov/housing/hvs/index.html)

Rental and homeowner vacancy rates and homeownership rates for the U.S., regions, states, and the 75 largest Metropolitan Statistical Areas. (Localities not included.) Updated quarterly.

## Data sources

[**American Community Survey**](https://www.census.gov/programs-surveys/acs) (ACS)  

The American Community Survey is an ongoing survey program of the U.S. Census Bureau. The ACS collects information about the nation’s social, economic, housing, and demographic characteristics on an annual basis. This information provides the most-up-to-date estimates between decennial censuses.

This section uses data from two ACS tables:

* **Table B25003**: Tenure
* **Table B25004**: Vacancy Status

The most recent data used for this section are the **2016-2020 5-year estimates**. These estimates average all survey responses across this period to increase sample sizes and reduce margins of error. Data from the previous 5-year estimates are also used as a baseline to show change over time.

```
^^^ EVERYTHING ABOVE GOES ON PAGE ^^^
```
