---
title: "Sourcebook > Rental > Rental vacancy"
subtitle: "Data sources and methods"
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output:
  html_document:
    theme: lumen
    toc: true
    toc_float: true
    number_sections: true
---
# Data sources

### American Community Survey (ACS) {-}

The American Community Survey is an ongoing survey program of the U.S. Census Bureau. The ACS collects information about the nation’s social, economic, housing, and demographic characteristics on an annual basis. This information provides the most-up-to-date estimates between decennial censuses.

This section uses data from two ACS tables:

* **Table B25003**: Tenure
* **Table B25004**: Vacancy Status

The most recent data used for this section are the **2015-2019 5-year estimates**. These estimates average all survey responses across this period to increase sample sizes and reduce margins of error.

Link: https://www.census.gov/programs-surveys/acs

# Setup

```{r setup, message=FALSE, warning=FALSE}

# Load packages and set global chunk options
library(tidycensus)
library(tidyverse)
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	eval = FALSE
)
```

# Data collection

```{r data-collect}

# Create an objects for years.

years <- 2010:2019

# Get variables for Table B25003

b25003_vars <- load_variables(2019, "acs5") %>%
  filter(str_detect(name, "B25003"))

# Get B25003 data for every locality in Virginia

b25003_raw <- map_dfr(years, function(yr){
  acs_pull <- get_acs(
    geography = "county",
    table = "B25003",
    year = yr,
    state = "VA"
  )
  
  acs_cleaned <- acs_pull %>%
    mutate(year = yr) %>%
    select(variable, year, NAME, GEOID,
           estimate, moe)
  
  acs_cleaned
})

# Get variables for Table B25004

b25004_vars <- load_variables(2019, "acs5") %>%
  filter(str_detect(name, "B25004"))

# Get B25004 data for every locality in Virginia

b25004_raw <- map_dfr(years, function(yr){
  acs_pull <- get_acs(
    geography = "county",
    table = "B25004",
    year = yr,
    state = "VA"
  )
  
  acs_cleaned <- acs_pull %>%
    mutate(year = yr) %>%
    select(variable, year, NAME, GEOID,
           estimate, moe)
  
  acs_cleaned
})

```

# Data prep
```{r data-prep}

# Clean B25003 variable names

b25003_vars_clean <- b25003_vars %>%
  separate(label, into = c("est", "tot", "tenure"),
           sep = "!!") %>%
  select(variable = name, tenure)%>%
  mutate(across(.fns = ~str_remove_all(.x, ":"))) %>%
  filter(across(c(tenure), ~!is.na(.x))) %>%
  filter(str_detect(tenure, "Owner occupied") == FALSE) %>% # Keep  only renter housing unit estimates
  mutate(tenure = case_when(
    str_detect(tenure, "Renter occupied") ~ "Renter",
    TRUE ~ tenure)
  )

# Join B25003 variables to data, calculate new sums, and add data reliability info

b25003_data <- b25003_raw %>%
  right_join(b25003_vars_clean, by = "variable") %>% # Use right_join to leave out unnecessary subtotals from data
  select(GEOID, year, tenure, estimate, moe) %>% # Simplify columns
  group_by(GEOID, year, tenure) %>% # Collapse and summarize
  summarise(
    estimate = sum(estimate),
    moe = moe_sum(moe, estimate) # Calculate new group-wise margins of error
  ) %>%
  ungroup()%>%
  mutate(cv = ((moe/1.645)/estimate)*100) %>% # Calculate coefficient of variation
  mutate(reliability = case_when(
    cv < 15 ~ "High reliability",
    cv >= 15 & cv <= 30 ~ "Medium reliability",
    cv > 30 ~ "Low reliability")
  )

# Clean B25004 variable names

b25004_vars_clean <- b25004_vars %>%
  separate(label, into = c("est", "total", "status"),
           sep = "!!") %>%
  select(variable = name, status) %>%
  mutate(across(.fns = ~str_remove_all(.x, ":"))) %>%
  filter(across(c(status), ~!is.na(.x))) %>%
  filter(grepl("rent", status, ignore.case = TRUE))

# Join B25004 variables to data, calculate new sums, and add data reliability info

b25004_data <- b25004_raw %>%
  right_join(b25004_vars_clean, by = "variable") %>% # Use right_join to leave out unnecessary subtotals from data
  select(GEOID, year, status, estimate, moe) %>% # Simplify columns
  group_by(GEOID, year, status) %>% # Collapse and summarize
  summarise(
    estimate = sum(estimate),
    moe = moe_sum(moe, estimate) # Calculate new group-wise margins of error
  ) %>%
  ungroup()%>%
   mutate(cv = ((moe/1.645)/estimate)*100) %>% # Calculate coefficient of variation
  mutate(reliability = case_when(
    cv < 15 ~ "High reliability",
    cv >= 15 & cv <= 30 ~ "Medium reliability",
    cv > 30 ~ "Low reliability")
  )

b25004_data_wide <- b25004_data %>%
  pivot_wider(
    names_from = status,
    values_from = c(estimate, moe, cv, reliability)
  )

rental_vacancy <- merge(b25003_data, b25004_data_wide, by = c("GEOID", "year"))

output_vacancy <- rental_vacancy %>%
  select(fips = GEOID, year, renter_occupied = estimate, for_rent_vacant = `estimate_For rent`, rented_vacant = `estimate_Rented, not occupied`) %>%
  mutate(total_units = select(., renter_occupied:rented_vacant) %>%
           rowSums(na.rm = TRUE))
  
```

# Data export
```{r data-export}

write_csv(output_vacancy, "data/rentervacancy.csv")
```

# Dashboard development

Saved for future dashboard development via Shiny.