---
title: "Sourcebook > Rental > Rental vacancy"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```

```
VVV EVERYTHING BELOW GOES ON PAGE VVV
```
The dashboards on this page show the rental vacancy across Virginia from 2010 to 
2019. Rental vacancy rates are a common indicator of rental demand in a community. 
Although there are [different measures of rental vacancy](https://www.stessa.com/blog/vacancy-rate/),
a general measure is the share of vacant rental units versus the total share of 
rental units in an area (both vacant and occupied). A low rental vacancy rate can 
indicate a strong demand for rental housing but limited supply can also exacerbate 
affordability challenges, while a high rental vacancy rate can indicate a weak rental 
market.

## Rental vacancy rate (American Community Survey)

The following dashboards shows the rental vacancy rate based on American Community 
Survey data for all Virginia localities from 2010 to 2019. This measure focuses on 
the physical vacancy rate. This is the share of vacant rental units (units that are 
not currently occupied) versus the total number of rental units (both vacant and occupied).

Data is available at the state, CBSA, and locality level.
<br>
<br>
```{r acs-vacancy}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-RentalvacancyrateACS/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

## Additional resources

*[What Vacancy Rates Tell You About a Housing Shortage (And What They Don't)](https://www.strongtowns.org/journal/2020/8/30/what-vacancy-rates-tell-you-about-a-housing-shortage)
*[Investopedia: Vacancy Rate](https://www.investopedia.com/terms/v/vacancy-rate.asp)

## Data sources

[**American Community Survey**](https://www.census.gov/programs-surveys/acs) (ACS)  

The American Community Survey is an ongoing survey program of the U.S. Census Bureau. The ACS collects information about the nation’s social, economic, housing, and demographic characteristics on an annual basis. This information provides the most-up-to-date estimates between decennial censuses.

This section uses data from two ACS tables:

* **Table B25003**: Tenure
* **Table B25004**: Vacancy Status

The most recent data used for this section are the **2015-2019 5-year estimates**. These estimates average all survey responses across this period to increase sample sizes and reduce margins of error.




```
^^^ EVERYTHING ABOVE GOES ON PAGE ^^^
```
