---
title: "Sourcebook > Homeownership > Homeownership rate"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```

```
VVV EVERYTHING BELOW GOES ON PAGE VVV
```

# Homeownership rate

The dashboards on this page include data on the homeownership rate of Virginia by
geography, race, ethnicity, and age group. 

<br>
<br>

## Homeownership rate map

This dashboard shows the homeownership rate across Virginia localities and regions from 2010 to
2020, as well as the overall homeownership in Virginia during this timeframe. Clicking individual localities on the map allows you to see their homeownership rate over time and at the selected year.

```{r homeownership}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Homeowershipratemap/ho_map_db?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

## Homeownership rate by race and ethnicity

This dashboard shows the homeownership rate across Virginia based on race and ethnicity 
from 2010 to 2020. Homeownership is a key factor in addressing the wealth gap that exists
between householders of color and their white counterparts. The following dashboard helps 
understand the persistent homeownership gap that disproportionately impacts households of 
color.

```{r homeownership-race}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Homeowershipgap/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

## Homeownership rate by age

This dashboard shows the homeownership rate across Virginia age group from 2010 to 
2020. For millennials, homeownership is increasingly becoming difficult to attain due 
to rising home prices coupled with stagnant wages and student debt. 

```{r homeownership-age}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Homeownershipratebyage/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

## Additional resources

**Brookings Institution**

[*Homeownership, racial segregation, and policy solutions to racial wealth equity*](https://www.brookings.edu/essay/homeownership-racial-segregation-and-policies-for-racial-wealth-equity/)

**HUD Office of Policy Development & Research (PD&R)**

[*Closing the African American Homeownership Gap*](https://www.huduser.gov/portal/pdredge/pdr-edge-featd-article-032221.html)

**Federal Reserve Bank of Richmond**

[*Student Debt vs. Homeownership*](https://www.richmondfed.org/publications/research/econ_focus/2017/q4/around_the_fed)

[*The Black Homeownership Gap in the Fifth District: How Does Your County Compare?*](https://www.richmondfed.org/publications/community_development/5th_district_footprint/2018/footprint_201810)

**Federal Reserve Bank of St. Louis**

[*Narrowing the Racial Homeownership Gap*](https://www.stlouisfed.org/open-vault/2021/november/narrowing-the-racial-homeownership-gap)

**Forbes**

[*U.S. Millennials: Home Ownership And The Growing Chasm Between Aspiration And Reality*](https://www.forbes.com/sites/nigelwilson/2021/08/18/us-millennials-home-ownership-and-the-growing-chasm-between-aspiration-and-reality/?sh=1e8288fc5498)

**Freddie Mac**

[*Millennials and Housing: Homeownership Demographic Research* PDF](https://sf.freddiemac.com/content/_assets/resources/pdf/fact-sheet/millennial-playbook_millennials-and-housing.pdf)

**Urban Institute**

[*Millennial Homeownership: Why Is It So Low, and How Can We Increase It?*](https://www.urban.org/research/publication/millennial-homeownership)

[*Reducing the Racial Homeownership Gap*](https://www.urban.org/policy-centers/housing-finance-policy-center/projects/reducing-racial-homeownership-gap)

## Data sources

[**American Community Survey**](https://www.census.gov/programs-surveys/acs) (ACS)

The American Community Survey is an ongoing survey program of the U.S. Census Bureau. The ACS collects information about the nation’s social, economic, housing, and demographic characteristics on an annual basis. This information provides the most-up-to-date estimates between decennial censuses.

This section uses data from two ACS tables:

* **Table B25003**: Tenure (iterations by race and ethnicity)
* **Table B25010**: Tenure by Age of Householder

The most recent data used for this section are the **2016-2020 5-year estimates**. These estimates average all survey responses across this period to increase sample sizes and reduce margins of error. Data from the previous 5-year estimates are also used as a baseline to show change over time.



```
^^^ EVERYTHING ABOVE GOES ON PAGE ^^^
```