---
title: "Sourcebook > Homeownership > Homeownership rate"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```

```
VVV EVERYTHING BELOW GOES ON PAGE VVV
```

# Homeownership rate

The dashboards on this page include data on the homeownership rate of Virginia by
locality, race, and ethnicity. 

```
[add stock photo?]
```
<br>
<br>

## Homeownership rate map

This dashboard shows the homeownership rate across Virginia localities from 2010 to
2019, as well as the overall homeownership in Virginia during this timeframe. Clicking
on individual localities allows you to see their homeownership rate over time and at
the selected year.

```{r homeownership}
knitr::include_url(url = "https://public.tableau.com/app/profile/housingforward.virginia/viz/Sourcebook-Homeowershipratemap/ho_map_db?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

## Homeownership rate by race and ethnicity

This dashboard shows the homeownership rate across Virginia based on race and ethnicity 
from 2010 to 2019. Homeownership is a key factor in addressing the wealth gap that exists
between householders of color and their white counterparts. The following dashboard helps 
understand the persistent homeownership gap that disproportionately impacts households of 
color.

```{r homeownership-race}
knitr::include_url(url = "https://public.tableau.com/app/profile/housingforward.virginia/viz/Sourcebook-Homeowershipgap/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

## Homeownership rate by age

This dashboard shows the homeownership rate across Virginia age group from 2010 to 
2019. For millennials, homeownership is increasingly becoming difficult to attain due 
to rising home prices coupled with stagnant wages and student debt. 

```{r homeownership-age}
knitr::include_url(url = "https://public.tableau.com/app/profile/housingforward.virginia/viz/Sourcebook-Homeownershipratebyage/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>
<br>

## Additional resources

**Brookings Institution**

[*Homeownership, racial segregation, and policy solutions to racial wealth equity*](https://www.brookings.edu/essay/homeownership-racial-segregation-and-policies-for-racial-wealth-equity/)

**HUD Office of Policy Development & Research (PD&R)**

[*Closing the African American Homeownership Gap*](https://www.huduser.gov/portal/pdredge/pdr-edge-featd-article-032221.html)

**Federal Reserve Bank of Richmond**

[*Student Debt vs. Homeownership*](https://www.richmondfed.org/publications/research/econ_focus/2017/q4/around_the_fed)

[*The Black Homeownership Gap in the Fifth District: How Does Your County Compare?*](https://www.richmondfed.org/publications/community_development/5th_district_footprint/2018/footprint_201810)

**Federal Reserve Bank of St. Louis**

[*Narrowing the Racial Homeownership Gap*](https://www.stlouisfed.org/open-vault/2021/november/narrowing-the-racial-homeownership-gap)

**Forbes**

[*U.S. Millennials: Home Ownership And The Growing Chasm Between Aspiration And Reality*](https://www.forbes.com/sites/nigelwilson/2021/08/18/us-millennials-home-ownership-and-the-growing-chasm-between-aspiration-and-reality/?sh=1e8288fc5498)

**Freddie Mac**
[*Millennials and Housing: Homeownership Demographic Research* PDF](https://sf.freddiemac.com/content/_assets/resources/pdf/fact-sheet/millennial-playbook_millennials-and-housing.pdf)

**Urban Institute**

[*Millennial Homeownership: Why Is It So Low, and How Can We Increase It?*](https://www.urban.org/research/publication/millennial-homeownership)

[*Reducing the Racial Homeownership Gap*](https://www.urban.org/policy-centers/housing-finance-policy-center/projects/reducing-racial-homeownership-gap)
<br>
<br>
<br>

## Data sources

[**American Community Survey**](https://www.census.gov/programs-surveys/acs) (ACS)
Period estimates based on the previous Decennial Census released annually. 


```
^^^ EVERYTHING ABOVE GOES ON PAGE ^^^
```