---
title: "Sourcebook > Rental > Rent"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```

```
VVV EVERYTHING BELOW GOES ON PAGE VVV
```
The dashboards on this page include data on median gross rent and the distribution of gross rent for renter-occupied housing units in Virginia from 2010 to 2020.

In the American Community Survey, the Census Bureau defines gross rent as the contract rent plus the estimated average monthly cost of utilities such as electricity, gas, water/sewer, and heating fuels such as oil or kerosene. 

## Median gross rent

The median gross rent helps determine the typical rent value in a community. Using the median prevents the effects that outliers, like a few high-end luxury rental units, may have on an average rent. 

Data is available at the state, CBSA, and locality level.
<br><br>

```{r med-rent}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Mediangrossrent/Statewide?:showVizHome=no&:embed=true", height = "500px")
```

<br><br>
To accurately compare dollar values across time, values must be converted to a consistent dollar value. This is because economic trends change the dollar's value over time. One hundred dollars five years ago is not the same as \$100 today due to [inflationary and deflationary](https://www.vox.com/2014/7/24/18080392/inflation-definition-and-explanation) forces.

We adjust the dollar value of median rent to the year in which data is most recently available, using the [Consumer Price Index: All Items Less Shelter](https://www.bls.gov/news.release/cpi.t03.htm) published by the U.S. Bureau of Labor Statistics (BLS).
<br><br>
 
## Rent distribution

Communities benefit from having an assortment of rental options to meet the needs of households of all incomes. The distribution of gross rent can help show whether a community has a balanced number of rental units at varying prices, or if a community is starting to skew towards a high-end rental market. 

Data is available at the state, CBSA, and locality level.
<br><br>

```{r rent-dist}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Grossrentdistribution/Statewide?:showVizHome=no&:embed=true", height = "500px")
```

<br><br>

## Additional resources

**Joint Center for Housing Studies of Harvard University**

[*Rental Housing*](https://www.jchs.harvard.edu/research-areas/rental-housing)

**Zumper**

[*Zumper's 2020 Annual Rent Report*](https://www.zumper.com/blog/zumper-2021-annual-rent-report/)

## Data sources

[**American Community Survey**](https://www.census.gov/programs-surveys/acs) (ACS)

The American Community Survey is an ongoing survey program of the U.S. Census Bureau. The ACS collects information about the nation’s social, economic, housing, and demographic characteristics on an annual basis. This information provides the most-up-to-date estimates between decennial censuses.

This section uses data from two ACS tables:

* **Table B24064**: Median Gross Rent (Dollars)
* **Table B25063**: Gross Rent

The most recent data used for this section are the **2016-2020 5-year estimates**. These estimates average all survey responses across this period to increase sample sizes and reduce margins of error. Data from the previous 5-year estimates are also used as a baseline to show change over time.

[**Federal Reserve Economic Data**](https://fred.stlouisfed.org/series/CUUR0000SA0L2) (FRED)

While the BLS officially calculates and publishes CPI values monthly, the Federal Reserve Bank of St. Louis also makes the data available through [FRED](https://fredhelp.stlouisfed.org/fred/about/about-fred/what-is-fred/). Historical CPI values, along with many other economic indicators, are easier to view and download in FRED.

```
^^^ EVERYTHING ABOVE GOES ON PAGE ^^^
```