---
title: "Sourcebook > Rental > Rent"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```

```
VVV EVERYTHING BELOW GOES ON PAGE VVV
```

## Rent

The dashboards on this page include data on median gross rent and the distribution of
gross rent for renter-occupied housing units in Virginia from 2010 to 2019. 

```
[add stock photo?]
```

### Median gross rent

Gross rent is defined as contract rent plus the estimated average monthly cost of
utilities such as electricity, gas, and water/sewer and fuels such as oil or kerosene. 

The median gross rent helps determine the typical rent value in a community. Using the
median prevents the affects that outliers, like a few high-end luxury rental units, may 
have on an average rent. 

Data is available at the state, CBSA, and locality level.

```{r med-rent}
knitr::include_url(url = "https://public.tableau.com/app/profile/housingforward.virginia/viz/Sourcebook-Mediangrossrent/Statewide?:showVizHome=no&:embed=true", height = "500px")
```

```
[insert blurb about comparing dollar values across time]
```

### Rent distribution

Communities benefit from having an assortment of rental options to meet the need of 
households of all incomes. The distribution of gross rent can help show whether a 
community has a balanced number of rental units at varying prices, or if a community 
is starting to skew towards a high-end rental market. 

Data is available at the state, CBSA, and locality level.

```{r rent-dist}
knitr::include_url(url = "https://public.tableau.com/app/profile/housingforward.virginia/viz/Sourcebook-Grossrentdistribution/Statewide?:showVizHome=no&:embed=true", height = "500px")
```

## Data sources


## Additional resources

```
[add relevant links]
```

```
^^^ EVERYTHING ABOVE GOES ON PAGE ^^^
```