---
title: "Sourcebook > Rental > Rent"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```

```
VVV EVERYTHING BELOW GOES ON PAGE VVV
```
The dashboards on this page include data on median gross rent and the distribution of
gross rent for renter-occupied housing units in Virginia from 2010 to 2020. 


## Median gross rent

Gross rent is defined as contract rent plus the estimated average monthly cost of
utilities such as electricity, gas, and water/sewer and fuels such as oil or kerosene. 

The median gross rent helps determine the typical rent value in a community. Using the
median prevents the affects that outliers, like a few high-end luxury rental units, may 
have on an average rent. 

Data is available at the state, CBSA, and locality level.
<br><br>
```{r med-rent}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Mediangrossrent/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br><br>
To accurately compare dollar values across time, values need to be converted to a consistent dollar value. This is because the value of the dollar changes across time. One hundred dollars five years ago is not the same as \$100 today due to changes in the price of goods, [inflation and deflation](https://www.vox.com/2014/7/24/18080392/inflation-definition-and-explanation).

We adjust the dollar value of median rent to the year in which data is most recently available utilizing the [Consumer Price Index: All Items Less Shelter](https://fred.stlouisfed.org/series/CUUR0000SA0L2).
<br><br>
 
## Rent distribution

Communities benefit from having an assortment of rental options to meet the need of 
households of all incomes. The distribution of gross rent can help show whether a 
community has a balanced number of rental units at varying prices, or if a community 
is starting to skew towards a high-end rental market. 

Data is available at the state, CBSA, and locality level.
<br><br>
```{r rent-dist}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Grossrentdistribution/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br><br>

## Additional resources

**Joint Center for Housing Studies of Harvard University**

[*Rental Housing*](https://www.jchs.harvard.edu/research-areas/rental-housing)

**Zumper**

[*Zumper's 2020 Annual Rent Report*](https://www.zumper.com/blog/zumper-2021-annual-rent-report/)


## Data sources

[*American Community Survey*](https://www.census.gov/programs-surveys/acs) (ACS

The American Community Survey is an ongoing survey program of the U.S. Census Bureau. The ACS collects information about the nation’s social, economic, housing, and demographic characteristics on an annual basis. This information provides the most-up-to-date estimates between decennial censuses.

```
^^^ EVERYTHING ABOVE GOES ON PAGE ^^^
```