---
title: "Sourcebook > Demographics > Age"
subtitle: "Data sources and methods"
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output:
  html_document:
    theme: lumen
    toc: true
    toc_float: true
    number_sections: true
---

# Data sources

### Census Population Estimates Program (PEP) {-}

Annual estimates of total population released by the U.S. Census Bureau. Available at locality level from 2010 to 2019. Estimates by age, race, and ethnicity also available, along with components of population change.

Link: https://www.census.gov/programs-surveys/popest.html

# Setup

```{r setup, message=FALSE, warning=FALSE}

# Load packages and set global chunk options

library(tidycensus)
library(tidyverse)

knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)

```

# Data collection

```{r collect}

# Download population counts by age group from PEP for each Virginia locality from 2010 to 2022
# 

age_vars <- c(
  "Age 0 to 4 years",
  "Age 5 to 9 years",
  "Age 10 to 14 years",
  "14 to 17 years",
  "18 to 24 years",
  "Age 25 to 29 years",
  "Age 30 to 34 years",
  "Age 35 to 39 years",
  "Age 40 to 44 years",
  "Age 45 to 49 years",
  "Age 50 to 54 years",
  "Age 55 to 59 years",
  "Age 60 to 64 years",
  "Age 65 to 69 years",
  "Age 70 to 74 years",
  "Age 75 to 79 years",
  "Age 80 to 84 years",
  "Age 85 years and older",
  "Median age"
)


age_raw_1 <- get_estimates(
  geography = "county",
  state = "VA",
  product = "characteristics",
  breakdown = "AGEGROUP",
  breakdown_labels = TRUE,
  year = 2019,
  time_series = TRUE
) %>%
  filter(!DATE %in% c(2, 3), # Remove non-decennial 2010 counts
         AGEGROUP %in% age_vars) %>% # Filter only selected age group variables
  mutate(counttype = # Add descriptions to count types
      case_when(
        DATE == 1 ~ "Census population",
        TRUE ~ "Population estimate")) %>%
  mutate(year = # Translate date codes into years
    case_when(
      DATE == 1 ~ "2010",
      DATE == 4 ~ "2011",
      DATE == 5 ~ "2012",
      DATE == 6 ~ "2013",
      DATE == 7 ~ "2014",
      DATE == 8 ~ "2015",
      DATE == 9 ~ "2016",
      DATE == 10 ~ "2017",
      DATE == 11 ~ "2018",
      DATE == 12 ~ "2019")) %>% 
  mutate(agegroup = # Recode detailed age groups to fewer ranges
    case_when(
      AGEGROUP == "Age 0 to 4 years" ~ "Under 10",
      AGEGROUP == "Age 5 to 9 years" ~ "Under 10",
      AGEGROUP == "Age 10 to 14 years" ~ "10 to 17",
      AGEGROUP == "14 to 17 years" ~ "10 to 17",
      AGEGROUP == "18 to 24 years" ~ "18 to 24",
      AGEGROUP == "Age 25 to 29 years" ~ "25 to 29",
      AGEGROUP == "Age 30 to 34 years" ~ "30 to 34",
      AGEGROUP == "Age 35 to 39 years" ~ "35 to 44",
      AGEGROUP == "Age 40 to 44 years" ~ "35 to 44",
      AGEGROUP == "Age 45 to 49 years" ~ "45 to 54",
      AGEGROUP == "Age 50 to 54 years" ~ "45 to 54",
      AGEGROUP == "Age 55 to 59 years" ~ "55 to 64",
      AGEGROUP == "Age 60 to 64 years" ~ "55 to 64",
      AGEGROUP == "Age 65 to 69 years" ~ "65 to 74",
      AGEGROUP == "Age 70 to 74 years" ~ "65 to 74",
      AGEGROUP == "Age 75 to 79 years" ~ "75 and over",
      AGEGROUP == "Age 80 to 84 years" ~ "75 and over",
      AGEGROUP == "Age 85 years and older" ~ "75 and over",
      AGEGROUP == "Median age" ~ "Median age"
    )) %>% 
  select( # Simplify columns
    GEOID,
    year,
    counttype,
    agegroup,
    value
  )


age_raw_2 <- get_estimates(
  geography = "county",
  state = "VA",
  product = "characteristics",
  breakdown = "AGEGROUP",
  breakdown_labels = TRUE,
  year = 2022,
  time_series = TRUE
) %>% 
  mutate(counttype = "Population estimate") |> 
  filter(AGEGROUP %in% age_vars) %>% # Filter only selected age group variables)
  mutate(agegroup = # Recode detailed age groups to fewer ranges
    case_when(
      AGEGROUP == "Age 0 to 4 years" ~ "Under 10",
      AGEGROUP == "Age 5 to 9 years" ~ "Under 10",
      AGEGROUP == "Age 10 to 14 years" ~ "10 to 17",
      AGEGROUP == "14 to 17 years" ~ "10 to 17",
      AGEGROUP == "18 to 24 years" ~ "18 to 24",
      AGEGROUP == "Age 25 to 29 years" ~ "25 to 29",
      AGEGROUP == "Age 30 to 34 years" ~ "30 to 34",
      AGEGROUP == "Age 35 to 39 years" ~ "35 to 44",
      AGEGROUP == "Age 40 to 44 years" ~ "35 to 44",
      AGEGROUP == "Age 45 to 49 years" ~ "45 to 54",
      AGEGROUP == "Age 50 to 54 years" ~ "45 to 54",
      AGEGROUP == "Age 55 to 59 years" ~ "55 to 64",
      AGEGROUP == "Age 60 to 64 years" ~ "55 to 64",
      AGEGROUP == "Age 65 to 69 years" ~ "65 to 74",
      AGEGROUP == "Age 70 to 74 years" ~ "65 to 74",
      AGEGROUP == "Age 75 to 79 years" ~ "75 and over",
      AGEGROUP == "Age 80 to 84 years" ~ "75 and over",
      AGEGROUP == "Age 85 years and older" ~ "75 and over",
      AGEGROUP == "Median age" ~ "Median age"
    )) %>% 
  select( # Simplify columns
    GEOID,
    year,
    counttype,
    agegroup,
    value
  )


age_raw <- rbind(age_raw_1, age_raw_2)

```

# Data prep

```{r prep}

# Create list of age variables to extract from data

# age_vars <- c(
#   "Age 0 to 4 years",
#   "Age 5 to 9 years",
#   "Age 10 to 14 years",
#   "14 to 17 years",
#   "18 to 24 years",
#   "Age 25 to 29 years",
#   "Age 30 to 34 years",
#   "Age 35 to 39 years",
#   "Age 40 to 44 years",
#   "Age 45 to 49 years",
#   "Age 50 to 54 years",
#   "Age 55 to 59 years",
#   "Age 60 to 64 years",
#   "Age 65 to 69 years",
#   "Age 70 to 74 years",
#   "Age 75 to 79 years",
#   "Age 80 to 84 years",
#   "Age 85 years and older",
#   "Median age"
# )
# 
# # Prep total population counts from PEP
# 
# age_clean <- age_raw %>%
#   filter(!DATE %in% c(2, 3), # Remove non-decennial 2010 counts
#          AGEGROUP %in% age_vars) %>% # Filter only selected age group variables
#   mutate(counttype = # Add descriptions to count types
#       case_when(
#         DATE == 1 ~ "Census population",
#         TRUE ~ "Population estimate")) %>%
#   mutate(year = # Translate date codes into years
#     case_when(
#       DATE == 1 ~ "2010",
#       DATE == 4 ~ "2011",
#       DATE == 5 ~ "2012",
#       DATE == 6 ~ "2013",
#       DATE == 7 ~ "2014",
#       DATE == 8 ~ "2015",
#       DATE == 9 ~ "2016",
#       DATE == 10 ~ "2017",
#       DATE == 11 ~ "2018",
#       DATE == 12 ~ "2019")) %>% 
#   mutate(agegroup = # Recode detailed age groups to fewer ranges
#     case_when(
#       AGEGROUP == "Age 0 to 4 years" ~ "Under 10",
#       AGEGROUP == "Age 5 to 9 years" ~ "Under 10",
#       AGEGROUP == "Age 10 to 14 years" ~ "10 to 17",
#       AGEGROUP == "14 to 17 years" ~ "10 to 17",
#       AGEGROUP == "18 to 24 years" ~ "18 to 24",
#       AGEGROUP == "Age 25 to 29 years" ~ "25 to 29",
#       AGEGROUP == "Age 30 to 34 years" ~ "30 to 34",
#       AGEGROUP == "Age 35 to 39 years" ~ "35 to 44",
#       AGEGROUP == "Age 40 to 44 years" ~ "35 to 44",
#       AGEGROUP == "Age 45 to 49 years" ~ "45 to 54",
#       AGEGROUP == "Age 50 to 54 years" ~ "45 to 54",
#       AGEGROUP == "Age 55 to 59 years" ~ "55 to 64",
#       AGEGROUP == "Age 60 to 64 years" ~ "55 to 64",
#       AGEGROUP == "Age 65 to 69 years" ~ "65 to 74",
#       AGEGROUP == "Age 70 to 74 years" ~ "65 to 74",
#       AGEGROUP == "Age 75 to 79 years" ~ "75 and over",
#       AGEGROUP == "Age 80 to 84 years" ~ "75 and over",
#       AGEGROUP == "Age 85 years and older" ~ "75 and over",
#       AGEGROUP == "Median age" ~ "Median age"
#     )) %>% 
#   select( # Simplify columns
#     GEOID,
#     year,
#     counttype,
#     agegroup,
#     value
#   )

age_data <- age_raw%>%
  group_by(GEOID, year, counttype, agegroup) %>% # Collapse and sum recoded age groups
  summarise(value = sum(value)) %>% 
  ungroup()

```

# Data export

```{r export}

# Save race and ethnicity data

write_csv(age_data, "data/age.csv")

```

# Dashboard development

Reserved for future Shiny app development.