---
title: "Sourcebook > Demographics > Age"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```

```
VVV EVERYTHING BELOW GOES ON PAGE VVV
```

## Age

The dashboards on this page break down population data by age groups. The amounts and relative shares of children, young adults, middle-age adults, and seniors within a community all significantly impact its different housing needs.
<br>
<br>
```
[add stock photo?]
```

### Population by age group

This dashboard shows the current population broken down into ten age groups. This data is from the 2019 release of the Census Bureau's Population Estimates Program. Total population counts by age groups from the 2020 Census will be available later in 2022.

Data is available at the state, CBSA, and locality level.
<br>
<br>
```{r pop-total}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Currentpopulationbyagegroup/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

### Change in population by age group

This dashboard shows changes in population by age group from 2010 to 2019. Data for 2010 is from that year's Decennial Census counts; all other years use estimates from the Census Bureau’s Population Estimates Program.

Data is available at the state, CBSA, and locality level.
<br>
<br>
```{r pop-components}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Changeinpopulationbyagegroup/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

## Data sources

[**Census Population Estimates Program**](https://www.census.gov/programs-surveys/popest.html) (PEP)  
Annual estimates of total population released by the U.S. Census Bureau. Available at locality level from 2010 to 2019. Estimates by age, race, and ethnicity also available, along with components of population change.

## Additional resources

```
[add relevant links]
```

```
^^^ EVERYTHING ABOVE GOES ON PAGE ^^^
```