---
title: "Sourcebook > Demographics > Household types"
subtitle: "Data sources and methods"
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output:
  html_document:
    theme: lumen
    toc: true
    toc_float: true
    number_sections: true
---

# Data sources

### American Community Survey (ACS) {-}

The American Community Survey is an ongoing survey program of the U.S. Census Bureau. The ACS collects information about the nation’s social, economic, housing, and demographic characteristics on an annual basis. This information provides the most-up-to-date estimates between decennial censuses.

This section uses data from two ACS tables:

* **Table B11012**: Households by Type
* **Table B09021**: Living Arrangements of Adults 18 Years and Over by Age

The most recent data used for this section are the **2015-2019 5-year estimates**. These estimates average all survey responses across this period to increase sample sizes and reduce margins of error.

Link: https://www.census.gov/programs-surveys/acs

# Setup

```{r setup, message=FALSE, warning=FALSE}

# Load packages and set global chunk options

library(tidycensus)
library(tidyverse)

knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

# Data collection

```{r collect}

# Get variables for Table B11012

b11012_vars <- load_variables(2019, "acs5") %>% 
  filter(str_detect(name, "B11012"))

# Get B11012 data for every locality in Virginia

b11012_raw <- get_acs(
  geography = "county",
  state = "VA",
  table = "B11012",
  year = 2019
)

# Get variables for Table B09021

b09021_vars <- load_variables(2019, "acs5") %>% 
  filter(str_detect(name, "B09021"))

# Get B09021 data for every locality in Virginia

b09021_raw <- get_acs(
  geography = "county",
  state = "VA",
  table = "B09021",
  year = 2019
)

```

# Data prep

```{r prep}

# Clean B11012 variable names

b11012_vars_clean <- b11012_vars %>% 
  separate(label, into = c("est", "tot", "type", "subtype"),
           sep = "!!") %>% 
  select(variable = name, type, subtype) %>% 
  mutate(across(.fns = ~str_remove_all(.x, ":"))) %>%
  filter(across(c(type, subtype), ~ !is.na(.x))) %>% 
  mutate(type = case_when(
    str_detect(type, "couple") ~ "Married or cohabitating couple",
    str_detect(type, "present") ~ "Householder with no partner",
    TRUE ~ type)
    ) %>% 
  mutate(subtype = case_when(
    str_detect(subtype, "With own children") ~ "With own children",
    str_detect(subtype, "With no own") ~ "Without own children",
    subtype == "With relatives, no own children under 18 years" ~ "With relatives, no own children",
    subtype == "With only non relatives present" ~ "With only nonrelatives",
    TRUE ~ subtype)
    )

# Join B11012 variables to data, calculate new sums, and add data reliability info

b11012_data <- b11012_raw %>% 
  right_join(b11012_vars_clean, by = "variable") %>% # Use right_join to leave out unnecessary subtotals from data
  select(GEOID, type, subtype, estimate, moe) %>% # Simplify columns
  group_by(GEOID, type, subtype) %>% # Collapse and sum subtypes
  summarise(
    estimate = sum(estimate),
    moe = moe_sum(moe, estimate) # Calculate new group-wise margins of error 
  ) %>% 
  ungroup() %>% 
  mutate(cv = ((moe/1.645)/estimate)*100) %>% # Calculate coefficient of variation
  mutate(reliability = case_when(
    cv < 15 ~ "High reliability",
    cv >= 15 & cv <= 30 ~ "Medium reliability",
    cv > 30 ~ "Low reliability")
  )

# Clean B09021 variable names

b09021_vars_clean <- b09021_vars %>% 
  separate(label, into = c("est", "tot", "age", "type"),
           sep = "!!") %>% 
  select(variable = name, age, type) %>% 
  mutate(across(.fns = ~str_remove_all(.x, ":")),
         age = str_remove_all(age, " years")) %>%
  mutate(type = case_when(
    variable %in% c(paste0("B09021_00", 2:7)) ~ age,
    TRUE ~ type)) %>%
  mutate(age = case_when(
    variable %in% c(paste0("B09021_00", 2:7)) ~ "All ages",
    TRUE ~ age)) %>%
  filter(across(c(age, type), ~ !is.na(.x))) %>% 
  mutate(type = case_when(
    str_detect(type, "Householder living") ~ "Lives with married or unmarried partner",
    type == "Child of householder" ~ "Lives with parent(s)",
    type == "Other relatives" ~ "Lives with other relative(s)",
    type == "Other nonrelatives" ~ "Lives with other nonrelative(s)",
    TRUE ~ type)
  )

# Join B09021 variables to data, calculate new sums, and add data reliability info

b09021_data <- b09021_raw %>% 
  right_join(b09021_vars_clean, by = "variable") %>% # Use right_join to leave out unnecessary subtotals from data
  select(GEOID, age, type, estimate, moe) %>% # Simplify columns
  group_by(GEOID, age, type) %>% # Collapse and sum subtypes
  summarise(
    estimate = sum(estimate),
    moe = moe_sum(moe, estimate) # Calculate new group-wise margins of error 
  ) %>% 
  ungroup() %>% 
  mutate(cv = ((moe/1.645)/estimate)*100) %>% # Calculate coefficient of variation
  mutate(reliability = case_when(
    cv < 15 ~ "High reliability",
    cv >= 15 & cv <= 30 ~ "Medium reliability",
    cv > 30 ~ "Low reliability")
  )

```

# Data export

```{r export}

# Save household types data from B11012

write_csv(b11012_data, "data/hh-type.csv")

# Save living arrangements data from B09021

write_csv(b09021_data, "data/living-arrangements.csv")

```

# Dashboard development

Reserved for future Shiny app development.