---
title: "Sourcebook > Demographics > Household types"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```

```
VVV EVERYTHING BELOW GOES ON PAGE VVV
```

## Household types

The dashboards on this page break down the population based on the types of households they live in and the people they share their homes with. These data are helpful in understanding the different types, sizes, and locations of homes that a community's population may need.
<br>
<br>
```
[add stock photo?]
```

### Households by type

This dashboard shows households by type, depending on whether the householder lives with a partner or not, and whether children, other relatives, or nonrelatives are present. These categories include:

* **Married or cohabitating couples**, by whether they live...
  + *With their own children*, or
  + *Without their own children*, because they do not have any, or because they are grown and live elsewhere;
* **Householders with no partners**, by whether they live...
  + *With their own children*,
  + *With relatives but none of their own children*, because they do not have any, or because they are grown and live elsewhere,
  + *With only nonrelatives present*, such as friends or roommates, or
  + *Alone*, with no one else.

Data is available at the state, CBSA, and locality level.
<br>
<br>
```{r pop-total}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Householdsbytype/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

### Living arrangements of adults

This dashboard shows the types of persons adults over the age of 18 share their homes with, including partners, parents, other relatives, and nonrelatives. This data does not consider the presence of any children the adult may have.

Data is available at the state, CBSA, and locality level.
<br>
<br>
```{r pop-components}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Livingarrangementsofadults/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

## Data sources

[**American Community Survey**](https://www.census.gov/programs-surveys/acs) (ACS)  

The American Community Survey is an ongoing survey program of the U.S. Census Bureau. The ACS collects information about the nation’s social, economic, housing, and demographic characteristics on an annual basis. This information provides the most-up-to-date estimates between decennial censuses.

This section uses data from two ACS tables:

* **Table B11012**: Households by Type
* **Table B09021**: Living Arrangements of Adults 18 Years and Over by Age

The most recent data used for this section are the **2015-2019 5-year estimates**. These estimates average all survey responses across this period to increase sample sizes and reduce margins of error.

## Additional resources

```
[add relevant links]
```

```
^^^ EVERYTHING ABOVE GOES ON PAGE ^^^
```