---
title: "Sourcebook > Demographics > Population"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```

```
VVV EVERYTHING BELOW GOES ON PAGE VVV
```
The dashboards on this page include data on total population counts and components of population change. These trends help reveal how much, and for what reasons, population levels may be increasing or decreasing over time.
<br>
<br>

## Total population

This dashboard shows annual population counts and percent changes from 2010 to 2020. Data for 2010 and 2020 are from those years' Decennial Census counts; all other years use estimates from the Census Bureau's Population Estimates Program (PEP).

Data is available at the state, CBSA, and locality level.
<br>
<br>
```{r pop-total}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Totalpopulation/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>
The 2020 Census was greatly [impacted by the pandemic](https://www.pbs.org/newshour/nation/how-the-coronavirus-pandemic-has-affected-the-2020-census), as well as the use of [differential privacy](https://demographics.coopercenter.org/census2020-dp). The Demographics Research Group at the University of Virginia's Weldon Cooper Center [advises data users to treat 2020 Census numbers with caution](https://statchatva.org/2021/08/27/why-2020-census-data-needs-to-be-treated-with-caution/). But overall, they find that 2020 Census data "tends to be more reliable for larger areas and populations whereas in Virginia’s smaller localities, data errors and distortions are much more common."
<br>

## Components of population change

This dashboard breaks down annual population changes by three different components:

* Natural increases (births minus deaths),
* Domestic migration (net moves from in/out of state), and
* International migration (net moves from in/out of the country).

Data is available at the state, CBSA, and locality level.
<br>
<br>
```{r pop-components}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Componentsofpopulationchange/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

## Additional resources

**U.S. Census Bureau**

[*Virginia adds more than 600,000 people since 2010*](https://www.census.gov/library/stories/state-by-state/virginia-population-change-between-census-decade.html)

**UVA Weldon Cooper Center**

[*State population growth slows, with most new residents in Northern Virginia*](https://news.virginia.edu/content/state-population-growth-slows-most-new-residents-northern-virginia)

[*Comparing 2020 Virginia population projections to the Census count*](https://statchatva.org/2021/09/27/comparing-2020-virginia-population-projections-to-the-census-count/)

## Data sources

[Census Population Estimates Program](https://www.census.gov/programs-surveys/popest.html) (PEP)

Annual estimates of total population released by the U.S. Census Bureau. Available at locality level from 2010 to 2019. Estimates by age, race, and ethnicity also available, along with components of population change.

[2020 Decennial Census](https://www.census.gov/programs-surveys/decennial-census/about/rdo/summary-files.html)

Initial data from the 2020 Decennial Census is available in the PL 94-171 Redistricting summary file. As of February 2022, we use this data to append 2020 population counts onto the PEP time series data through 2019. Complete 2010 to 2020 Intercensal Estimates will be published in late 2022.



```
^^^ EVERYTHING ABOVE GOES ON PAGE ^^^
```