---
title: "Sourcebook > Demographics > Population"
subtitle: "Data sources and methods"
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output:
  html_document:
    theme: lumen
    toc: true
    toc_float: true
    number_sections: true
---

# Data sources

### Census Population Estimates Program (PEP) {-}

Annual estimates of total population released by the U.S. Census Bureau. Available at locality level from 2010 to 2019. Estimates by age, race, and ethnicity also available, along with components of population change.

Link: https://www.census.gov/programs-surveys/popest.html

### 2020 Decennial Census {-}

Initial data from the 2020 Decennial Census is available in the PL 94-171 Redistricting summary file. As of February 2022, we use this data to append 2020 population counts onto the PEP time series data through 2019. Complete 2010 to 2020 Intercensal Estimates will be published in late 2022.

Link: https://www.census.gov/programs-surveys/decennial-census/about/rdo/summary-files.html

### Weldon Cooper Center for Public Service {-}

Official population projections produced by the Cooper Center Demographics Research Group at the University of Virginia. Current projections available at the locality level for 2030 and 2040. New projections based on 2020 Census (for 2030 to 2050) expected by July 1, 2022.

Link: https://demographics.coopercenter.org/virginia-population-projections

# Setup

```{r setup, message=FALSE, warning=FALSE}

# Load packages and set global chunk options

library(tidycensus)
library(tidyverse)
library(knitr)
library(janitor)

knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)

```

# Data collection

```{r pop-collect}

# Download total population counts from PEP for each Virginia locality from 2010 to 2019

pep_total_raw <- get_estimates(
  geography = "county",
  state = "VA",
  variables = "POP",
  year = 2019,
  time_series = TRUE
)

# Download components of population change from PEP for each Virginia locality from 2010 to 2019

pep_change_raw <- get_estimates(
  geography = "county",
  state = "VA",
  variables = c("NATURALINC", "DOMESTICMIG", "INTERNATIONALMIG"),
  year = 2019,
  time_series = TRUE
)

# Download total population counts from 2020 Census summary file for each Virginia locality

census_raw <- get_decennial(
  geography = "county",
  state = "VA",
  year = 2020,
  sumfile = "pl",
  variables = "P1_001N"
)

```

# Data prep

```{r pop-prep}

# Prep total population counts from PEP

pep_total_clean <- pep_total_raw %>%
  filter(!DATE %in% c(2, 3)) %>% # Remove non-decennial 2010 counts
  mutate(year = # Translate date codes into years
    case_when(
      DATE == 1 ~ "2010",
      DATE == 4 ~ "2011",
      DATE == 5 ~ "2012",
      DATE == 6 ~ "2013",
      DATE == 7 ~ "2014",
      DATE == 8 ~ "2015",
      DATE == 9 ~ "2016",
      DATE == 10 ~ "2017",
      DATE == 11 ~ "2018",
      DATE == 12 ~ "2019")) %>% 
  mutate(counttype = # Add descriptions to count types
      case_when(
        DATE == 1 ~ "Census population",
        TRUE ~ "Population estimate")) %>% 
  select( # Simplify columns
    GEOID,
    year,
    counttype,
    value
  )

# Prep total population counts from 2020 Census summary file

census_clean <- census_raw %>% 
  mutate(year = "2020", # Add year and count type columns
         counttype = "Census population") %>%
  select( # Simplify columns
    GEOID,
    year,
    counttype,
    value
  )

# Append 2020 Census counts to PEP time series

population_data <- pep_total_clean %>% 
  bind_rows(census_clean)

# Prep components of population change from PEP

pep_change_clean <- pep_change_raw %>% 
  mutate(year = # Translate date codes into years
    case_when(
      PERIOD == 1 ~ "2010",
      PERIOD == 2 ~ "2011",
      PERIOD == 3 ~ "2012",
      PERIOD == 4 ~ "2013",
      PERIOD == 5 ~ "2014",
      PERIOD == 6 ~ "2015",
      PERIOD == 7 ~ "2016",
      PERIOD == 8 ~ "2017",
      PERIOD == 9 ~ "2018",
      PERIOD == 10 ~ "2019")) %>%
  mutate(component = # Rename components of change
    case_when(
      variable == "NATURALINC" ~ "Natural increase",
      variable == "DOMESTICMIG" ~ "Domestic migration",
      variable == "INTERNATIONALMIG" ~ "International migration")) %>% 
  select( # Simplify columns
    GEOID,
    year,
    component,
    value
  )

```

# Data export

```{r pop-export}

# Save total population counts

write_csv(population_data, "data/pop-total.csv")

# Save components of population change

write_csv(pep_change_clean, "data/pop-components.csv")

```

# Dashboard development

Reserved for future Shiny app development.