---
title: "Sourcebook > Affordability > Renter Affordability Match"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```

```
VVV EVERYTHING BELOW GOES ON PAGE VVV
```
CHAS data provide a tabulation of renter household income and whether the price of 
housing matches that income. With this data, we are able to get an estimate of how 
many renters are living in housing that is affordable to them. 

This data can be used to provide an approximation of the rental housing gap based on 
household income, in other words, the number of rental units that are needed at a 
specific affordability level.

Data is available at the state, CBSA, and locality level.
<br>
<br>
```{r rent-mismatch}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Rentalhousinggap/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>


## Data sources
[Comprehensive Housing Affordability Strategy (CHAS)](https://www.huduser.gov/portal/datasets/cp.html)

CHAS data are sourced from custom tabulations of the American Community Survey (ACS) 5-year estimates and are provided to the U.S. Department of Housing and Urban Development. Data are available at the locality level from 2009 to 2018. 

Inconsistencies exist across data dictionaries before 2012. Therefore, data is only presented from 2012-2018, when data variables are consistent.

