---
title: "Sourcebook > Affordability > Renter Affordability Match"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```

```
VVV EVERYTHING BELOW GOES ON PAGE VVV
```
The U.S. Census Bureau provides the U.S. Department of Housing and Urban Development (HUD) with tabulations of renter household incomes by area median income (AMI), and whether the current price of housing matches those incomes. With this data, we are able to get an estimate of how many renters are living in homes that are affordable (or unaffordable) to them. 

This data can be used to provide an approximation of the rental housing gap based on household income. In other words, this is the number of rental units needed at a specific affordability level to fully "eliminate" cost burden.

Data is available at the state, CBSA, and locality level.
<br>
<br>
```{r rent-mismatch}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Rentalhousinggap/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

## Additional resources

**National Low Income Housing Coalition**

[*The Gap: A Shortage of Affordable Homes*](https://nlihc.org/gap)

## Data sources

[Comprehensive Housing Affordability Strategy](https://www.huduser.gov/portal/datasets/cp.html) (CHAS)

CHAS data are sourced from custom tabulations of the American Community Survey (ACS) 5-year estimates and are provided to the U.S. Department of Housing and Urban Development. Data are available at the locality level from 2009 to 2018.

Inconsistencies exist across data dictionaries before 2012. Therefore, Sourcebook only uses CHAS data from 2012 to 2018, when data variables are consistent.

```
^^^ EVERYTHING ABOVE GOES ON PAGE ^^^
```