---
title: "Sourcebook > Demographics > Race and ethnicity"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```

```
VVV EVERYTHING BELOW GOES ON PAGE VVV
```
The dashboards on this page break down population data by race and ethnicity. This information helps advocates and practitioners understand demographic shifts in their community to effectively orient policy and programs for housing and other related services.
<br>
<br>
For an explanation of racial and ethnicity terms, please visti our SOURCEBOOK data dictionary [here](https://housingforwardva.org/toolkits/sourcebook/data-dictionary/).

## Current population by race and ethnicity

This dashboard shows population by race and ethnicity as of the 2020 Census. These are total counts and are therefore the most accurate representation of a community's diversity currently available.

Data is available at the state, CBSA, and locality level.
<br>
<br>
```{r race-current}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Currentpopulationbyraceandethnicity/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>
The 2020 Census was greatly [impacted by the pandemic](https://www.pbs.org/newshour/nation/how-the-coronavirus-pandemic-has-affected-the-2020-census), as well as the use of [differential privacy](https://demographics.coopercenter.org/census2020-dp). The Demographics Research Group at the University of Virginia's Weldon Cooper Center [advises data users to treat 2020 Census numbers with caution](https://statchatva.org/2021/08/27/why-2020-census-data-needs-to-be-treated-with-caution/). But overall, they find that 2020 Census data "tends to be more reliable for larger areas and populations whereas in Virginia’s smaller localities, data errors and distortions are much more common."
<br>

## Population change by race and ethnicity

This dashboard shows changes in population by race and ethnicity from 2010 to 2020. Data for 2010 and 2020 are from those years’ Decennial Census counts; all other years use estimates from the Census Bureau’s Population Estimates Program (PEP).

Data is available at the state, CBSA, and locality level.
<br>
<br>
```{r race-change}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Populationchangebyraceandethnicity/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

The greater-than-expected rise in the non-white population—and commensurate shrinking of the white population—surprised researchers when the 2020 Census results were released.

This phenomenon is very likely in part because of the [changing ways](https://www.npr.org/2021/08/22/1029609786/2020-census-data-results-white-population-shrinking-decline-non-hispanic-race) Americans are choosing to identify their race and ethnicity.
<br>

## Additional resources

**Brookings Institute**

[*The nation is diversifying even faster than predicted, according to new census data*](https://www.brookings.edu/research/new-census-data-shows-the-nation-is-diversifying-even-faster-than-predicted/)

**U.S. Census Bureau**

[*2020 U.S. population more racially and ethnically diverse than measured in 2010*](https://www.census.gov/library/stories/2021/08/2020-united-states-population-more-racially-ethnically-diverse-than-2010.html)

## Data sources

[**Census Population Estimates Program**](https://www.census.gov/programs-surveys/popest.html) (PEP)  
Annual estimates of total population released by the U.S. Census Bureau. Available at locality level from 2010 to 2019. Estimates by age, race, and ethnicity also available, along with components of population change.

[**2020 Decennial Census**](https://www.census.gov/programs-surveys/decennial-census/about/rdo/summary-files.html)  
Initial data from the 2020 Decennial Census is available in the PL 94-171 Redistricting summary file. As of February 2022, we use this data to append 2020 population counts onto the PEP time series data through 2019. Complete 2010 to 2020 Intercensal Estimates will be published in late 2022.



```
^^^ EVERYTHING ABOVE GOES ON PAGE ^^^
```