---
title: "Sourcebook > Demographics > Race and ethnicity"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```

```
VVV EVERYTHING BELOW GOES ON PAGE VVV
```

## Race and ethnicity

The dashboards on this page break down population data by race and ethnicity. This information helps advocates and practitioners understand demographic shifts in their community to effectively orient policy and programs for housing and other related services.
<br>
<br>
```
[add link to Sourcebook reference page on race/ethnicity terms]
```

```
[add stock photo?]
```

### Current population by race and ethnicity

This dashboard shows population by race and ethnicity as of the 2020 Census. These are total counts and are therefore the most accurate representation of a community's diversity currently available.

Data is available at the state, CBSA, and locality level.
<br>
<br>
```{r race-current}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Currentpopulationbyraceandethnicity/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>
```
[insert cautionary blurb on 2020 Census]
```
### Population change by race and ethnicity

This dashboard shows changes in population by race and ethnicity from 2010 to 2020. Data for 2010 and 2020 are from those years’ Decennial Census counts; all other years use estimates from the Census Bureau’s Population Estimates Program (PEP).

Data is available at the state, CBSA, and locality level.
<br>
<br>
```{r race-change}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Populationchangebyraceandethnicity/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>
```
[insert cautionary blurb on interpreting race and ethnicity changes in the 2020 Census]
```
## Additional resources

**Brookings Institute**

[*The nation is diversifying even faster than predicted, according to new census data*](https://www.brookings.edu/research/new-census-data-shows-the-nation-is-diversifying-even-faster-than-predicted/)

**U.S. Census Bureau**

[*2020 U.S. population more racially and ethnically diverse than measured in 2010*](https://www.census.gov/library/stories/2021/08/2020-united-states-population-more-racially-ethnically-diverse-than-2010.html)

## Data sources

[**Census Population Estimates Program**](https://www.census.gov/programs-surveys/popest.html) (PEP)  
Annual estimates of total population released by the U.S. Census Bureau. Available at locality level from 2010 to 2019. Estimates by age, race, and ethnicity also available, along with components of population change.

[**2020 Decennial Census**](https://www.census.gov/programs-surveys/decennial-census/about/rdo/summary-files.html)  
Initial data from the 2020 Decennial Census is available in the PL 94-171 Redistricting summary file. As of February 2022, we use this data to append 2020 population counts onto the PEP time series data through 2019. Complete 2010 to 2020 Intercensal Estimates will be published in late 2022.



```
^^^ EVERYTHING ABOVE GOES ON PAGE ^^^
```