---
title: "Sourcebook > Demographics > Race and Ethnicity"
subtitle: "Data sources and methods"
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output:
  html_document:
    theme: lumen
    toc: true
    toc_float: true
    number_sections: true
---

# Data sources

### Census Population Estimates Program (PEP) {-}

Annual estimates of total population released by the U.S. Census Bureau. Available at locality level from 2010 to 2019. Estimates by age, race, and ethnicity also available, along with components of population change.

Link: https://www.census.gov/programs-surveys/popest.html

### 2020 Decennial Census {-}

Initial data from the 2020 Decennial Census is available in the PL 94-171 Redistricting summary file. As of February 2022, we use this data to append 2020 population counts onto the PEP time series data through 2019. Complete 2010 to 2020 Intercensal Estimates will be published in late 2022.

Link: https://www.census.gov/programs-surveys/decennial-census/about/rdo/summary-files.html

# Setup

```{r setup, message=FALSE, warning=FALSE}

# Load packages and set global chunk options

library(tidycensus)
library(tidyverse)

knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)

```

# Data collection

```{r collect}

# Download population counts by race and ethnicity from PEP for each Virginia locality from 2010 to 2019

race_ethnicity_raw <- get_estimates(
  geography = "county",
  state = "VA",
  product = "characteristics",
  breakdown = c("RACE", "HISP"),
  breakdown_labels = TRUE,
  year = 2019,
  time_series = TRUE
)

# Get race and ethnicity variables from the 2020 Census summary file

census_vars <- load_variables(2020, "pl")

census_vars_race <- c(
  "P1_001N", # All races
  "P1_003N", # White alone
  "P1_004N", # Black or African American alone
  "P1_005N", # American Indian and Alaska Native alone
  "P1_006N", # Asian alone
  "P1_007N", # Native Hawaiian and Other Pacific Islander alone
  "P1_008N", # Some Other Race alone
  "P1_009N", # Two or more races
  "P2_002N", # Total Hispanic or Latino
  "P2_005N", # Hispanic or Latino, White alone
  "P2_006N", # Hispanic or Latino, Black or African American alone
  "P2_007N", # Hispanic or Latino, American Indian and Alaska Native alone
  "P2_008N", # Hispanic or Latino, Asian alone
  "P2_009N", # Hispanic or Latino, Native Hawaiian and Other Pacific Islander alone
  "P2_010N", # Hispanic or Latino, Some Other Race alone
  "P2_011N", # Hispanic or Latino, Two or more races
  "P2_003N"  # Total Not Hispanic or Latino
  )

# Download population counts by race and ethnicity from 2020 Census summary file for each Virginia locality

census_raw <- get_decennial(
  geography = "county",
  state = "VA",
  year = 2020,
  sumfile = "pl",
  variables = census_vars_race
)

```

# Data prep

```{r prep}

# Prep total population counts from PEP

race_ethnicity_clean <- race_ethnicity_raw %>%
  filter(!DATE %in% c(2, 3)) %>% # Remove non-decennial 2010 counts
  filter(!str_detect(RACE, "combination")) %>% # Remove counts for "in combination" race categories for simplicity
  mutate(year = # Translate date codes into years
    case_when(
      DATE == 1 ~ "2010",
      DATE == 4 ~ "2011",
      DATE == 5 ~ "2012",
      DATE == 6 ~ "2013",
      DATE == 7 ~ "2014",
      DATE == 8 ~ "2015",
      DATE == 9 ~ "2016",
      DATE == 10 ~ "2017",
      DATE == 11 ~ "2018",
      DATE == 12 ~ "2019")) %>% 
  mutate(counttype = # Add descriptions to count types
      case_when(
        DATE == 1 ~ "Census population",
        TRUE ~ "Population estimate")) %>% 
  select( # Simplify columns
    GEOID,
    year,
    counttype,
    race = RACE,
    hisp = HISP,
    value
  )

# Create and add labels to race and ethnicity variables from 2020 Census summary file

census_labels_race <- census_vars %>% 
  filter(name %in% census_vars_race) %>%
  separate(label, into = c("col1", "col2", "col3", "col4", "col5"), sep = "!!") %>% 
  mutate(across(.fns = ~str_remove_all(.x, ":")),
         across(.fns = ~replace_na(.x, "All"))) %>% 
  mutate(
    col4 = case_when(
      col5 == "White alone" ~ col5,
      col5 == "Black or African American alone" ~ col5,
      col5 == "American Indian and Alaska Native alone" ~ col5,
      col5 == "Asian alone" ~ col5,
      col5 == "Native Hawaiian and Other Pacific Islander alone" ~ col5,
      col5 == "Some Other Race alone" ~ col5,
      col3 == "Population of two or more races" ~ "Population of two or more races",
      col4 == "All" ~ "All races",
      TRUE ~ col4),
    col3 = case_when(
      col3 == "Population of one race" ~ "Both Hispanic Origins",
      col3 == "Population of two or more races" ~ "Both Hispanic Origins",
      col3 == "All" ~ "Both Hispanic Origins",
      col3 == "Hispanic or Latino" ~ "Hispanic",
      col3 == "Not Hispanic or Latino" ~ "Non-Hispanic",
      TRUE ~ col3),
    across(.fns = ~str_replace_all(.x, "Population of two or more races", "Two or more races")),
    across(.fns = ~str_replace_all(.x, "Black or African American alone", "Black alone"))
    ) %>% 
  select(
    variable = name,
    race = col4,
    hisp = col3
  )

# Prep total population counts from 2020 Census summary file

census_clean <- census_raw %>%
  left_join(census_labels_race, # Add race and ethnicity labels
    by = "variable"
  ) %>%
  select(-c(variable, NAME)) %>% # Remove unnecessary columns
  pivot_wider( # Pivot out values by Hispanic status
    names_from = "hisp",
    values_from = "value"
    ) %>% 
  mutate( # Calculate Hispanic values for all races (not included in Census data)
    `Hispanic` = `Both Hispanic Origins` - `Non-Hispanic`
  ) %>% 
  pivot_longer( # Transform data back into tidy format
   cols = contains("Hispanic"),
   names_to = "hisp",
   values_to = "value"
  ) %>%
  mutate(year = "2020", # Add year and count type columns to match PEP data
         counttype = "Census population") %>%
  select( # Simplify columns
    GEOID,
    year,
    counttype,
    race,
    hisp,
    value
  )

# Append 2020 Census data to PEP time series, remove unnecessary subtotals, and relabel variables

race_ethnicity_data <- race_ethnicity_clean %>% 
  bind_rows(census_clean) %>% 
  mutate(race = str_remove_all(race, " alone")) %>% # Remove "alone" from race column values
  filter( # Remove unnecessary subtotals to avoid double-counted data
    !hisp == "Both Hispanic Origins",
    hisp == "Non-Hispanic" | race == "All races") %>%
  mutate(label = # Create "label" column that collapses less prevalent race categories
    case_when(
      race == "All races" & hisp == "Hispanic" ~ "Hispanic or Latino",
      race == "White" ~ "White, non-Hispanic",
      race == "American Indian and Alaska Native" ~ "Another race",
      race == "Native Hawaiian and Other Pacific Islander" ~ "Another race",
      race == "Two or more races" ~ "Multiracial",
      race == "Some Other Race" ~ "Another race",
      TRUE ~ race
    )) %>% 
  filter(!label == "All races") %>% # Remove unnecessary "All races, Non-Hispanic" subtotal
  select( # Reorder columns
    GEOID,
    year,
    counttype,
    race,
    hisp,
    label,
    value
  )
  
```

# Data export

```{r export}

# Save race and ethnicity data

write_csv(race_ethnicity_data, "data/race-ethnicity.csv")

```

# Dashboard development

Reserved for future Shiny app development.