---
title: "Sourcebook > Inventory > Housing age"
subtitle: HTML content for website
author: "HousingForward Virginia"
date: "Updated on `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      out.width = "100%",
                      out.extra = 'style="border:none;"')
```


```
VVV EVERYTHING BELOW GOES ON PAGE VVV
```

## Housing age

The dashboards on this page break down housing units by their age. This is represented by the reported year each home was built. These data are helpful in understanding historical production trends over time, as well as serving as a proxy for housing quality. Older homes generally require more maintenance to be safe and efficient.
<br>
<br>
```
[add stock photo?]
```

### Housing age by tenure

This dashboard shows the current supply of occupied housing units among owners, renters, and all households, broken down by the year each home was built. Years are grouped into ten categories: 1939 or earlier, eight decades from 1940 through 2009, 2010 to 2013, and 2014 or later. The last two groupings are likely to change in the future as new ACS estimates are released annually.

Data is available at the state, CBSA, and locality level.
<br>
<br>
```{r dashboard}
knitr::include_url(url = "https://public.tableau.com/views/Sourcebook-Housingagebytenure/Statewide?:showVizHome=no&:embed=true", height = "500px")
```
<br>
<br>

## Data sources

[**American Community Survey**](https://www.census.gov/programs-surveys/acs)

The American Community Survey is an ongoing survey program of the U.S. Census Bureau. The ACS collects information about the nation’s social, economic, housing, and demographic characteristics on an annual basis. This information provides the most-up-to-date estimates between decennial censuses.

This section uses data from one ACS table:

* **Table B25036**: Tenure by Year Structure Built

The most recent data used for this section are the **2015-2019 5-year estimates**. These estimates average all survey responses across this period to increase sample sizes and reduce margins of error.

## Additional resources

```
[add relevant links]
```

```
^^^ EVERYTHING ABOVE GOES ON PAGE ^^^
```